### 2016-03-03 tacl2015_paper/decision.tex
adding tacl2015 paper
Former-commit-id: 0d96938966b2ba3d3b01c8b2bbda93814689193d

\documentclass{article}[11pt,oneside]

\begin{document}

\begin{verbatim}

As TACL action editor for submission 548, "Learning a Compositional
Semantics for Freebase with an Open Predicate Vocabulary",   I am happy to
tell you that I am accepting your paper subject (conditional) to your making
specific revisions within two months.

LIST OF MANDATORY REVISIONS:
The following issues were particularly prominent in the reviews below:

* There is a recent paper that seems closely related, "Large-scale semantic
parsing without Question-Answer Pairs", TACL 2014, Reddy, Lapata, and
Steedman. Please do a thorough comparison to this.

* Please add more discussion on the coverage and expressivity of the
approach: Which language phenomena are addressed by the representation in
this paper? How is negation handled, and which cases of negation are covered
by the current set of rules?

* Please do a comparison of probabilistic databases to Markov Logic
Networks, and more generally to probabilistic logic with a probability
distribution over
possible worlds.

But please also take into account all other reviewer comments.

Note that the reviewers asked for changes to the explanations of your model,
but not any additional experiments.

Generally, your revised version will be handled by the same action editor
(me) and the same reviewers (if necessary) in making the final decision ---
which, *if* all requested revisions are made, will be final acceptance.

You are allowed one to two extra pages of content to accommodate these
revisions.  To submit your revised version, follow the instructions in the
"Revision and Resubmission Policy for TACL Submissions" section of the
Author Guidelines at
https://tacl2013.cs.columbia.edu/ojs/index.php/tacl/about/submissions#authorGuidelines
.

Thank you for submitting to TACL, and I look forward to your revised
version!

Katrin Erk
University of Texas, Austin
katrin.erk@mail.utexas.edu
------------------------------------------------------
------------------------------------------------------
....THE REVIEWS....
------------------------------------------------------
------------------------------------------------------
Reviewer A:

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        5. Very clear.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        4. Creative: An intriguing problem, technique, or approach that is
substantially different from previous research.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        4. Some of the ideas or results will substantially help other people's
ongoing research.

REPLICABILITY Will members of the ACL community be able to reproduce or
verify the results in this paper?
Members of the ACL community:
:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF ACCOMPANYING SOFTWARE: If software was submitted along with the
paper, what is the expected impact of the software package?:
        1. No usable software released.

IMPACT OF ACCOMPANYING DATASET: If a dataset was submitted along with the
paper, what is its expected impact? Will it be valuable to others in the
form in which they are released?:
        1. No usable datasets submitted.

RECOMMENDATION: Should the paper be accepted or rejected? In deciding on
your ultimate recommendation, please think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
reccommend it. If a paper is solid but you could live without it, let us
know that you're ambivalent.:
        4. Worthy: A good paper that is worthy of being published in TACL.

Detailed Comments for the Authors:
        This is one of the best papers in the recent literature on combining formal
and distributional semantics. The model maps CCG parses to logical forms,
then learns an embedding for the property and relation symbols based on
their denotations. These embeddings are used to construct a probabilistic
database, in a similar way to universal schema. The approach allows them to
capture compositional semantics whilst retaining the 'softness' of
distributional semantics.

Most of my comments are minor.

The paper introduces a new ranking function which allows compositional
expressions to be added to the database - which I think is an important
contribution. The paper claims that their model represents "arbitrary
natural language phrases" (end of section 2) - but later it suggests that
only expressions in which all variables are bound to Freebase entities are
added. For example, I think that this restriction means that a sentence like
"Every president is a person" wouldn't be added to the database (c.f.
Rochtaschel et al. 2014, who are interested in adding first-order knowledge
to probabilistic databases). The paper should be more clear about exactly
what knowledge is being represented in their database.

I also think that the paper would benefit from a discussion of the
strengths/weaknesses of using a probabilistic database for knowledge
representation, compared to alternatives such as Markov logic networks (as
used by Beltagy et al. 2013). For example, I think that adding a single new
fact to the database would require it to be retrained?

In the evaluation, they find that manually-annotated Freebase queries
outperform their model. One approach would be to also include Freebase
predicates in the database (as in universal schema), and see if the
combination outperforms Freebase alone. If this isn't possible, explain why.
Also, I'd be interested to see a result using parsed Freebase queries
(rather than gold queries), e.g. using Berant's parser.

A clustering baseline performs weakly on the task - it's perhaps worth
mentioning what kinds of errors it's making (e.g. over-clustering,
under-clustering, etc.) Lewis and Steedman (2013) use some additional
techniques, such as modeling ambiguity and dropping weak links in the
clustering, which may explain the difference in performance.

The authors suggest that recall could be improved by training on a larger
corpus. The amount of training data they use actually seems relatively small
by recent standards - is the bottleneck in syntactic parsing times, or in
the learning?

Some of the post-processing CCG parses sounds very similar to that used by
Lewis and Steedman (such as converting verbal adjuncts to arguments, adding
word features to prepositions). If so, this should be referenced.

The related work should discuss the recent Reddy, Lapata and Steedman paper
(TACL 2014). Like the submission, they learn a semantics for predicates in a
CCG-derived logical form based on their Freebase-entity arguments. There are
several important differences, but they're worth discussing.

The paper is generally very well-written, but I think Figure 1 could be made
more clear. Also, Figure 3 should use human-readable constants rather than
Freebase IDs.

REVIEWER CONFIDENCE:
        4. Quite sure. I tried to check the important points carefully. It's
unlikely, though conceivable, that I missed something that should affect my
ratings.

------------------------------------------------------

------------------------------------------------------
Reviewer B:

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        3. Mostly understandable to me with some effort.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        3. Respectable: A nice research contribution that represents a notable
extension of prior approaches or methodologies.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        4. Some of the ideas or results will substantially help other people's
ongoing research.

REPLICABILITY Will members of the ACL community be able to reproduce or
verify the results in this paper?
Members of the ACL community:
:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF ACCOMPANYING SOFTWARE: If software was submitted along with the
paper, what is the expected impact of the software package?:
        1. No usable software released.

IMPACT OF ACCOMPANYING DATASET: If a dataset was submitted along with the
paper, what is its expected impact? Will it be valuable to others in the
form in which they are released?:
        1. No usable datasets submitted.

RECOMMENDATION: Should the paper be accepted or rejected? In deciding on
your ultimate recommendation, please think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
reccommend it. If a paper is solid but you could live without it, let us
know that you're ambivalent.:
        4. Worthy: A good paper that is worthy of being published in TACL.

Detailed Comments for the Authors:
        Summary of paper: The goal of this paper is to extend compositional
semantic parsing to open-domain predicates, as not to be bound by knowledge
bases such as Freebase. Specifically, a huge number of sentences are parsed
into logical forms using a simple rule-based semantic parser, where the
predicates (categories and relations) used are words.  These words are
embedded into a vector space via matrix factorization (like in universal
schema).  Experiments demonstrate the value of using matrix factorization
and training on compositional utterances.

Summary of assessment: Overall, I thought this was a decent paper.  The goal
of extending semantic parsing to more realistic settings is definitely a
worthwhile one.  The paper presents a reasonable and natural approach, and
the experiments demonstrate some promise in this approach.  There are,
however, several improvements that could be made that would turn this into a
strong paper:

1) I'd like to see a more rigorous treatment of the semantics.  This paper
has the opporunity to lay out a foundation for how compositional semantics
should be handled, but it comes across currently as rather ad-hoc.  Several
things about the exposition could be improved, as detailed below.

2) At some level, I'm not convinced by the practical utility of learning
these matrix factorization models based on the particular restricted
compositional utterances used in the paper, because at the end of the day,
you're just trying to model a set of relations.  For example, the utterance
"Tom Cruise plays Maverick in the movie Top Gun" is really just "Tom Cruise
plays Maverick" and "Tom Cruise plays in the movie Top Gun", which are both
just triples.  More complex utterances "Top Gun played in 5 movies" couldn't
be represented with triples and would actually require semantic parsing, but
this is not handled in the paper.  I understand that there is a 1\%
improvement by learning from compositional parses, but there's no analysis
of this.

More detailed comments:

Figure 2: the treatment of appositive constructions seems inappropriate.
Rather than stringing together a huge conjunction, the appositive
construction should be an independent conventional implicature statement,
whose validity is independent of the main statement.

Page 5: when presenting the probabilistic denotations, it would be clearer
to define things more rigorously.  For example, does this corresponds to a
world model where each ground predicate is true with some probability
independently? I believe the answer is no, and that there's some
approximation in the recursive definition of denotations, which should be
made clearer.  Comparing with what one gets from Markov logic would be
helpful.

I would have liked the paper to be more explicit early on about the exact
set of logical forms handled in two senses: (i) the formal logical
restrictions (e.g., only conjunctions and existential quantification is
allowed), and (ii) the practical sizes of the logical forms (e.g., do they
include 5 predicates or 50 predicates)?  I know statistics are given later
(in Table 1), but I think it really helps the reader understand the
motivation and design decisions if this appeared earlier.

5.1: the description of extracting simplified logical forms is not clear to
me. The examples help, but I didn't get a sense of what the general recipe
is on an arbitrarily logical form.  This seems like an important part of the
paper.

Perhaps state in 5.2 that this only handles simple predicate-entity queries
(the third example in Figure 3)?  Also, is $t_j'$ sampled based on $t_j$ so
that they share one entity in common.  In general, this part should be
spelled out in more detail: random from what distribution?

I found the ranking objective in 5.3 to be confusing.  In particular, in the
definition of $C(\ell)$ and $R_L(\ell)$, and $R_R(\ell)$, it's not clear
what this set is and that $y$ stands in for a concrete entity (?) in $\ell$,
but that $x$ is a variable (which doesn't even appear in $\ell$)?  This
definition also doesn't seem to generalize to all logical forms $\ell$.
What if the entity $e_i$ was embedded deeply in some large logical form
$\ell$ with lots of other existentially quantified variables?  I don't see
how the construction in the paper works for that, or is that outside the
scope of the paper?  Either way, being explicit would be useful.

At a higher-level note, the ranking objective seems rather ad-hoc.  Since
you already have defined a probabilistic denotation for a logical form
$\ell$ (section 4.2), why not just maximize the probability that the
denotation $\ell_i$ contains $e_i$, and the probability that denotation
$\ell_i$ does not contain random $e_i'$ that get sampled (appropriately
weighting the two terms)? I suppose the gradient updates would be more
involved and require some backpropagation, but conceptually, this would be a
lot cleaner, and the treatment of probabilistic semantics would be handled
in one place rather than two.

I was surprised that the Freebase semantic parsing method actually has more
complete coverage given that the folklore assumption (and indeed the
motivation for relation extraction) is that there is more data in text.

I would have liked to see much more error analysis of concrete examples of
compositional utterances that that system gets right and wrong.  What do the
examples in the test set look like?

REVIEWER CONFIDENCE:
        4. Quite sure. I tried to check the important points carefully. It's
unlikely, though conceivable, that I missed something that should affect my
ratings.

------------------------------------------------------

------------------------------------------------------
Reviewer C:

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        5. Very clear.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        4. Creative: An intriguing problem, technique, or approach that is
substantially different from previous research.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        5. The approach is very apt, and the claims are convincingly supported.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        3. Bibliography and comparison are somewhat helpful, but it could be hard
for a reader to determine exactly how this work relates to previous work or
what its benefits and limitations are.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        4. Some of the ideas or results will substantially help other people's
ongoing research.

REPLICABILITY Will members of the ACL community be able to reproduce or
verify the results in this paper?
Members of the ACL community:
:
        4. could mostly reproduce the results, but there may be some
variation because of sample variance or minor variations in their
interpretation of the protocol or method.

IMPACT OF ACCOMPANYING SOFTWARE: If software was submitted along with the
paper, what is the expected impact of the software package?:
        1. No usable software released.

IMPACT OF ACCOMPANYING DATASET: If a dataset was submitted along with the
paper, what is its expected impact? Will it be valuable to others in the
form in which they are released?:
        1. No usable datasets submitted.

RECOMMENDATION: Should the paper be accepted or rejected? In deciding on
your ultimate recommendation, please think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
reccommend it. If a paper is solid but you could live without it, let us
know that you're ambivalent.:
        5. Strong: I'd like to see it accepted; it will be one of the better papers
in TACL.

Detailed Comments for the Authors:
        Beautifully written paper describing an extension to the Freebase semantic
parsing problem, incorporating matrix factorisation borrowed from the ideas
on universal schema. The authors have worked hard to compare against
ompetitive baselines, and argue convincingly for the open nature of the
represenation. Careful thought has been given to the machine learning side,
with extensions to previous objective functions used for similar work.

The only grumble I have relates to an apparent lack of awareness of a highly
relevant existing TACL paper. It is perhaps generous to suggest an oversight
on the part of the authors in not citing the following paper which came out
in October 2014:

Large-scale Semantic Parsing without Question-Answer Pairs
Siva Reddy, Mirella Lapata, Mark Steedman

The submitted paper must position itself with respect to this paper, since
both use CCG for freebase semantic parsing. (The Reddy paper does not use an
"open" representation, so I can see this crucial difference.)

More minor comments follow.

Try this piece of tex to get your CCG categories looking much cleaner
(without the extraneous space):

\newcommand{\cf}[1]{\mbox{$\it{#1}$}}   % category font, e.g. \cf{NP}

Modifiers of verbs are changed to complements using PP. What if there are
lots of modifiers? That implies lots of PPs? eg (...((S\NP)/PP/PP)...)?
Maybe this doesn't matter if the change is only done in post-processing
before the semantic rules are applied, but it might be a bad idea to have
this representation as part of the supertagger and/or parser.

Special rules are used to deal with negation. How is this dealt with in the
semantics. Eg, off the top of my head, what about a query like "All the
James Bond actors that did *not* appear in Bond movies before 1980?" Is it
possible to take set complements? Is that how it's handled?

Footnote 2 is a little opaque. I understand the difference between the
marginal and conditional prob (I think); something like P(entity|query) vs.
P(entity, query), so in the marginal case there's no distribution over
possible answers. But why one works better than the other is mysterious, and
footnote 2 didn't help much.

Just one typo (!): "as opposed tuples in general"

Section 6.2: "Each system predicts..." - which systems? It's not clear at
this point what the systems are providing the pool.

I'm also not entirely convinced by the use of the pooling methodology here.
My understanding of pooling in IR is that experiments have been run showing
that, with enough diversity in the systems making up the pool, the recall in
the pooled answers is extremely high, making them a good basis for manual
annotation (since no relevant documents or correct answers are missed). If
the systems you use are just the handful you evaluate, it's not clear
pooling has the same guarantees.

I was also puzzled by the idea that there can be more than one answer to a
query, so a couple of examples here would really help. (I guess there could
be more than one CEO of a company, for example?)

It also wasn't obvious to me that MAP is the right evaluation, if there are
so few possible answers. Why not just rank of the first correct answer? What
do they do in, eg, TREC QA track here?

The last line of table 1 (> 1 answer 116) confused me. Is this the number of
questions with more than one answer? In the text it says at least one (so
>=) *system* able to produce the correct answer.

lambda = 10^-4, 300 dimensions: were these parameters tuned? If so, on what
data? (I guess I'm wondering whether additional tuning was done for your
system compared to the baselines, and on what data...)

I would change the title of section 7 (Discussion) to Conclusion, since
there's very little discussion.

REVIEWER CONFIDENCE:
        4. Quite sure. I tried to check the important points carefully. It's
unlikely, though conceivable, that I missed something that should affect my
ratings.

\end{verbatim}

\end{document}