### 2016-03-03 tacl2015_paper/paper.tex

\documentclass[11pt]{article}

\usepackage{acl2012}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[pdf]{pstricks}
\usepackage{pst-pdf}
\usepackage{pst-plot}
\usepackage{times}
\usepackage{multirow}
\usepackage{latexsym}
\usepackage{soul}
\usepackage{subfig}
\usepackage{stmaryrd}
\usepackage{url}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

\title{Learning a Compositional Semantics for Freebase \\ with an Open
  Predicate Vocabulary}

\author{Jayant Krishnamurthy\\
        Carnegie Mellon University\\
        5000 Forbes Avenue\\
        Pittsburgh, PA 15213\\
	{\tt jayantk@cs.cmu.edu}
	  \And
	Tom M. Mitchell\\
        Carnegie Mellon University\\
        5000 Forbes Avenue\\
        Pittsburgh, PA 15213\\
        {\tt tom.mitchell@cmu.edu}}

\date{}

% Contributions:
% - compositional model of semantics based on universal schema (cite old work)
% - new training objective
% - experimental evaluation
% 
% Introduction:
% 
% - Fixed schema is inherently limiting
% - open / universal is good
%   - need techniques for composition / compositional semantics
% -

\newcommand{\pred}[1]{\textsc{#1}}
% TODO: make this look cleaner!
\newcommand{\ccgsyn}[1]{\mbox{$\it{#1}$}}
\newcommand{\true}{\pred{True}}

\newcommand{\clb}{\textsc{CorpusLookup}}
\newcommand{\cb}{\textsc{Clustering}}
\newcommand{\uspr}{\textsc{Factorization} (\ensuremath{O_P})}
\newcommand{\usqr}{\textsc{Factorization} (\ensuremath{O_Q})}
\newcommand{\epr}{\textsc{Ensemble} (\ensuremath{O_P})}
\newcommand{\eqr}{\textsc{Ensemble} (\ensuremath{O_Q})}
\newcommand{\pden}[1]{\ensuremath{\llbracket #1 \rrbracket_{P}}}

\DeclareMathOperator{\precision}{Prec}
\DeclareMathOperator{\correct}{Correct}

\begin{document}
\maketitle
\begin{abstract}
We present an approach to learning a model-theoretic semantics for
natural language tied to Freebase. Crucially, our approach uses an
open predicate vocabulary, enabling it to produce denotations for
phrases such as ``Republican front-runner from Texas'' whose semantics
cannot be represented using the Freebase schema. Our approach directly
converts a sentence's syntactic CCG parse into a logical form
containing predicates derived from the words in the sentence,
assigning each word a consistent semantics across sentences. This
logical form is evaluated against a learned probabilistic database
that defines a distribution over denotations for each textual
predicate. A training phase produces this probabilistic database using
a corpus of entity-linked text and probabilistic matrix factorization
with a novel ranking objective function. We evaluate our approach on a
compositional question answering task where it outperforms several
competitive baselines. We also compare our approach against manually
annotated Freebase queries, finding that our open predicate vocabulary
enables us to answer many questions that Freebase cannot.

% This approach allows us to learn denotations for tricky words (e.g.,
% ``front-runner'') and compose them to produce denotations for phrases
% (e.g., ``Republican front-runner from Texas''). The use of CCG also
% enables us to model linguistic phenomena such as negation. 
\end{abstract}

\section{Introduction}

Traditional knowledge representation assumes that world knowledge can
be encoded using a closed vocabulary of formal predicates. In recent
years, semantic parsing has enabled us to build compositional models
of natural language semantics using such a closed predicate vocabulary
\cite{zelle1996,zettlemoyer05}. These semantic parsers map natural
language statements to database queries, enabling applications such as
answering questions using a large knowledge base
\cite{yahya12,krishnamurthy2012,cai2013acl,kwiatkowski2013,berant2013,berant2014,reddy2014}. Furthermore,
the model-theoretic semantics provided by such parsers have the
potential to improve performance on other tasks, such as information
extraction and coreference resolution.

However, a closed predicate vocabulary has inherent
limitations. First, its coverage will be limited, as such vocabularies
are typically manually constructed. Second, it may abstract away
potentially relevant semantic differences. For example, the semantics
of ``Republican front-runner'' cannot be adequately encoded in the
Freebase schema because it lacks the concept of a ``front-runner.''
We could choose to encode this concept as ``politician'' at the cost
of abstracting away the distinction between the two. As this example
illustrates, these two problems are prevalent in even the largest
knowledge bases.

\begin{figure*}
\small
\begin{minipage}{0.23\linewidth}\textbf{Input Text}\\Republican front-runner \\ from Texas\end{minipage} ~~ $\rightarrow$ ~~
\begin{minipage}{0.3\linewidth}
\textbf{Logical Form}\\
$\lambda x.\exists y,z. \pred{front-runner}(x) \wedge y=\pred{/en/republican} \wedge \pred{nn}(y,x) \wedge z=\pred{/en/texas} \wedge \pred{from}(x,z)$
\end{minipage} ~~ $\rightarrow$ ~~
\begin{minipage}{0.3\linewidth}
\begin{tabular}{ll}
\textbf{Entity} & \textbf{Prob.} \\
$\pred{/en/george\_bush}$ & 0.57 \\
$\pred{/en/rick\_perry}$ & 0.45 \\
... & \\
\end{tabular}
\end{minipage}

\vspace{-.35in}
\begin{minipage}{0.15\linewidth}
\begin{pspicture}(-0.25,-0.4)(4.2,3.6)
\scalebox{0.8}{

\psframe(1,0)(2,2)
\psframe(2.2,1)(4.2,2)
\rput(1.5, 1){$\phi$}
\rput(3.2, 1.5){$\theta$}

\rput[rt](0.95,1.85){{\small \pred{ texas}}}
\rput[rt](0.95,1.35){{\small \pred{ repub.}}}
\rput[rt](0.95,0.85){{\small \pred{ g.\_bush}}}
\rput[rt](0.95,0.35){{\small \pred{ ...}}}

\rput[lb]{60}(2.4,2){{\small \pred{f.-run.}}}
\rput[lb]{60}(2.9,2){{\small \pred{pol.}}}
\rput[lb]{60}(3.4,2){{\small \pred{state}}}
\rput[lb]{60}(3.9,2){{\small \pred{...}}}

\rput[l](0,-0.3){\large \textbf{Entity/Predicate Embeddings}}
}
\end{pspicture}
\end{minipage}
\begin{minipage}{0.35\linewidth}
\begin{pspicture}(-2.7,-0.4)(4.7,3.6)
\scalebox{0.8}{

\psframe(1,0)(2,2)
\psframe(2.2,1)(4.2,2)
\rput(1.5, 1){$\phi$}
\rput(3.2, 1.5){$\theta$}

\rput[rt](0.95,1.85){{\small (\pred{g.\_bush}, \pred{ texas})}}
\rput[rt](0.95,1.35){{\small (\pred{g.\_bush}, \pred{ repub.})}}
\rput[rt](0.95,0.85){{\small (\pred{repub.}, \pred{ g.\_bush})}}
\rput[rt](0.95,0.35){{\small \pred{...}}}

\rput[lb]{60}(2.4,2){{\small \pred{from}}}
\rput[lb]{60}(2.9,2){{\small \pred{lives\_in}}}
\rput[lb]{60}(3.4,2){{\small \pred{nn}}}
\rput[lb]{60}(3.9,2){{\small \pred{...}}}

\rput(4.6,1.5){$\rightarrow$}
}
\end{pspicture}
\end{minipage}
\begin{minipage}{0.15\linewidth}
\begin{pspicture}(-1,-0.4)(8,3.6)
\scalebox{0.8}{

\psframe(1,0)(3,2)
\psline(1,0.5)(3,0.5)
\psline(1,1)(3,1)
\psline(1,1.5)(3,1.5)
\psline(1.5,0)(1.5,2)
\psline(2,0)(2,2)
\psline(2.5,0)(2.5,2)

\rput[rt](0.95,1.85){{\small \pred{ texas}}}
\rput[rt](0.95,1.35){{\small \pred{ repub.}}}
\rput[rt](0.95,0.85){{\small \pred{ g.\_bush}}}
\rput[rt](0.95,0.35){{\small \pred{ ...}}}

\rput[lb]{60}(1.3,2){{\small \pred{f.-run.}}}
\rput[lb]{60}(1.8,2){{\small \pred{pol.}}}
\rput[lb]{60}(2.3,2){{\small \pred{state}}}
\rput[lb]{60}(2.8,2){{\small \pred{...}}}

\rput[c](1.25,1.75){{\small .1}}
\rput[c](1.25,1.25){{\small .1}}
\rput[c](1.25,0.75){{\small .9}}

\rput[c](1.75,1.75){{\small .1}}
\rput[c](1.75,1.25){{\small .1}}
\rput[c](1.75,0.75){{\small .9}}

\rput[c](2.25,1.75){{\small .8}}
\rput[c](2.25,1.25){{\small .1}}
\rput[c](2.25,0.75){{\small .1}}

\rput[l](0,-0.3){\large \textbf{Probabilistic Database}}

% \psline[linewidth=0.5pt](3.4,1.5)(6.5,1.5)
% \psline[linewidth=0.5pt]{->}(6.5,1.5)(6.5,3.7)
}
\end{pspicture}
\end{minipage}
\begin{minipage}{0.2\linewidth}
\begin{pspicture}(-2.5,-0.4)(8,3.6)
\scalebox{0.8}{

\psframe(1,0)(3,2)
\psline(1,0.5)(3,0.5)
\psline(1,1)(3,1)
\psline(1,1.5)(3,1.5)
\psline(1.5,0)(1.5,2)
\psline(2,0)(2,2)
\psline(2.5,0)(2.5,2)

\rput[rt](0.95,1.85){{\small (\pred{g.\_bush}, \pred{ texas})}}
\rput[rt](0.95,1.35){{\small (\pred{g.\_bush}, \pred{ repub.})}}
\rput[rt](0.95,0.85){{\small (\pred{repub.}, \pred{ g.\_bush})}}
\rput[rt](0.95,0.35){{\small \pred{ ...}}}

\rput[lb]{60}(1.3,2){{\small \pred{from}}}
\rput[lb]{60}(1.8,2){{\small \pred{lives\_in}}}
\rput[lb]{60}(2.3,2){{\small \pred{nn}}}
\rput[lb]{60}(2.8,2){{\small \pred{...}}}

\rput[c](1.25,1.75){{\small .9}}
\rput[c](1.25,1.25){{\small .1}}
\rput[c](1.25,0.75){{\small .1}}

\rput[c](1.75,1.75){{\small .9}}
\rput[c](1.75,1.25){{\small .1}}
\rput[c](1.75,0.75){{\small .1}}

\rput[c](2.25,1.75){{\small .1}}
\rput[c](2.25,1.25){{\small .1}}
\rput[c](2.25,0.75){{\small .7}}

% \psline[linewidth=0.5pt](3.4,1.5)(6.5,1.5)
% \psline[linewidth=0.5pt]{->}(6.5,1.5)(6.5,3.7)
\psline[linewidth=0.5pt]{->}(-0.5,2.8)(-0.5,3.2)

}
\end{pspicture}
\end{minipage}
\vspace{-.1in}
\caption{Overview of our approach. Top left: the text is converted to
  logical form by CCG syntactic parsing and a collection of
  manually-defined rules. Bottom: low-dimensional embeddings of each
  entity (entity pair) and category (relation) are learned from an
  entity-linked web corpus. These embeddings are used to construct a
  probabilistic database. The labels of these matrices are shortened
  for space reasons. Top right: evaluating the logical form on the
  probabilistic database computes the marginal probability that each
  entity is an element of the text's denotation.}
\label{fig:overview}
\vspace{-.1in}
\end{figure*}

An alternative paradigm is an \emph{open} predicate vocabulary, where
each natural language word or phrase is given its own formal
predicate. This paradigm is embodied in both open information
extraction \cite{banko2007} and universal schema
\cite{riedel2013}. Open predicate vocabularies have the potential to
capture subtle semantic distinctions and achieve high
coverage. However, we have yet to develop compelling approaches to
compositional semantics within this paradigm.

% An alternative approach is \emph{open information extraction}, which
% treats every natural language word or phrase as its own formal
% predicate \cite{banko2007}. Another closely related approach is
% \emph{universal schema}, which joins these natural language predicates
% with formal predicates \cite{riedel2013}. 
% In these approaches, the knowledge representation has an open
% predicate vocabulary that is determined by the textual expressions in
% a corpus.

This paper takes a step toward compositional semantics with an open
predicate vocabulary. Our approach defines a distribution over
denotations (sets of Freebase entities) given an input text.  The
model has two components, shown in Figure \ref{fig:overview}. The
first component is a rule-based semantic parser that uses a syntactic
CCG parser and manually-defined rules to map entity-linked texts to
logical forms containing predicates derived from the words in the
text. The second component is a probabilistic database with a possible
worlds semantics that defines a distribution over denotations for each
textually-derived predicate. This database assigns independent
probabilities to individual predicate instances, such as
$P(\pred{front-runner}(\pred{/en/george\_bush})) = 0.9$. Together,
these components define an exponentially-large distribution over
denotations for an input text; to simplify this output, we compute the
marginal probability, over all possible worlds, that each entity is an
element of the text's denotation.

% A training phase estimates parameters for the probabilistic database
% using probabilistic matrix factorization with a novel query/answer
% ranking objective. The training data is derived by analyzing a corpus
% of entity-linked sentences with the rule-based semantic analysis
% system, producing a collection logical form queries with observed
% entity answers. The query/answer ranking objective, when optimized,
% trains the database to rank the observed answers for each query above
% unobserved answers. The matrix factorization learns a low-dimensional
% embedding of each entity (entity pair) and category (relation) such
% that the denotation of a predicate is likely to contain entities or
% entity pairs with nearby vectors.

The learning problem in our approach is to train the probabilistic
database to predict a denotation for each predicate. We pose this
problem as probabilistic matrix factorization with a novel
query/answer ranking objective. This factorization learns a
low-dimensional embedding of each entity (entity pair) and category
(relation) such that the denotation of a predicate is likely to
contain entities or entity pairs with nearby vectors. To train the
database, we first collect training data by analyzing entity-linked
sentences in a large web corpus with the rule-based semantic
parser. This process generates a collection of logical form queries
with observed entity answers. The query/answer ranking objective, when
optimized, trains the database to rank the observed answers for each
query above unobserved answers.

% Given a learned probabilistic database, Figure \ref{fig:overview}
% illustrates the process of predicting a denotation for a text. First,
% the semantic analysis system produces a logical form for the given
% text (top left). Next, this logical form is evaluated against the
% learned probabilistic database to produce a list of entities ranked by
% probability, similar to existing semantic parsing techniques (top
% right). Each predicate instance's probability in the database is
% determined by the low-dimensional embeddings learned by matrix
% factorization (bottom left).

% pose this problem as
% probabilistic matrix factorization with a novel ranking
% objective. First, we collect training data, in the form of true
% assertions about Freebase entities found in a large web corpus. We
% then automatically generate query-answer pairs from these assertions
% and train the database to rank  Training produces a vector embedding of each
% predicate and entity such that the denotation of a predicate is likely
% to contain entities with nearby vectors.
% 
% 
% An initial training phase uses the semantic analysis component and a
% large entity-linked corpus to learn parameters for the probabilistic
% database,  The set of categories and
% relations is automatically derived from the words in the training
% data.

% The denotation of a text is computed in three
% steps.  (Embeddings of entity pairs and relation
% predicates are also learned, but not shown in the Figure for space
% reasons.) Second, given a text, a rule based-semantic analysis system
% produces a logical form. 
% 
% Our approach represents world knowledge in a
% probabilistic database over textually-derived predicates and Freebase
% entities. To semantically analyze a natural language statement, we use a
% rule-based approach to transform its syntactic CCG parse into a query
% against this database. Evaluating the query produces a ranked list of
% Freebase entity answers,

% We train the database using a data set of true assertions about
% Freebase entities collected from a large web corpus. , or learning
% embeddings for each predicated and entity that assign high probability
% to the observed assertions.

%  Our approach first uses a CCG-based semantic analysis system
% to identify assertions about Freebase entities, resulting in a sparse
% collection of observed predicate instances. Our approach then
% generalizes from the observed predicate instances using probabilistic
% matrix factorization (TODO: cite). We introduce a novel training
% objective that encourages the learned parameters to correctly answer
% compositional natural language queries.

We evaluate our approach on a question answering task, finding that
our approach outperforms several baselines and that our new training
objective improves performance over a previously-proposed
objective. We also evaluate the trade-offs between open and closed
predicate vocabularies by comparing our approach to a
manually-annotated Freebase query for each question. This comparison
reveals that, when Freebase contains predicates that cover the
question, it achieves higher precision and recall than our
approach. However, our approach can correctly answer many questions
not covered by Freebase.

% The contributions of this work are threefold:
% \begin{enumerate} 
% \item An approach to compositional semantics with an open predicate
%   vocabulary that uses probabilistic matrix factorization to produce
%   the underlying knowledge representation.
% 
% \item A novel query/answer ranking objective for the matrix
%   factorization that improves performance over a previously-proposed
%   objective.
% 
% \item An experimental evaluation on a question answering task that
%   highlights the advantages and disadvantages of open and closed
%   predicate vocabularies.
% \end{enumerate}

% (TODO: key challenges to semantic analysis: (1) ability to represent
% phenomena like negation, etc. (2) synonymy)

% \section{Overview}
% 
% This paper is structured as follows. Section \ref{sec:rule} describes
% the rule-based semantic analysis system, which is used to generate
% training data for the probabilistic database. Section \ref{sec:probdb}
% describes the probabilistic database, our matrix factorization model
% and our training algorithm. Section \ref{sec:evaluation} describes our
% experimental evaluation, Section \ref{sec:priorwork} describes prior
% work, and Section \ref{sec:discussion} concludes.

\section{System Overview}

The purpose of our system is to predict a denotation $\gamma$ for a
given natural language text $s$. The denotation $\gamma$ is the set of
Freebase entities that $s$ refers to; for example, if
$s=\mbox{``president of the US,''}$ then $\gamma = \{\pred{/en/Obama},
\pred{/en/Bush}, ... \}$.\footnote{This paper uses a simple
  model-theoretic semantics where the denotation of a noun phrase is a
  set of entities and the denotation of a sentence is either true or
  false. However, for notational convenience, denotations $\gamma$
  will be treated as sets of entities throughout.} Our system
represents this prediction problem using the following probabilistic
model:
\begin{eqnarray*}
P(\gamma | s) = \sum_w \sum_\ell P(\gamma | \ell, w) P(w) P(\ell | s) 
\end{eqnarray*}

The first term in this factorization, $P(\ell | s)$, is a distribution
over logical forms $\ell$ given the text $s$. This term
corresponds to the rule-based semantic parser (Section
\ref{sec:rule}). This semantic parser is deterministic, so this term
assigns probability 1 to a single logical form for each text. The
second term, $P(w)$, represents a distribution over possible worlds,
where each world is an assignment of truth values to all possible
predicate instances. The distribution over worlds is represented by a
probabilistic database (Section \ref{sec:probdb}). The final term,
$P(\gamma | \ell, w)$, deterministically evaluates the logical form
$\ell$ on the world $w$ to produce a denotation $\gamma$. This term
represents query evaluation against a fixed database, as in other work
on semantic parsing.

Section \ref{sec:inference} describes inference in our model. To
produce a ranked list of entities (Figure \ref{fig:overview}, top
right) from $P(\gamma | s)$, our system computes the marginal
probability that each entity is an element of the denotation
$\gamma$. This problem corresponds to query evaluation in a
probabilistic database, which is known to be tractable in many cases
\cite{suciu2011probabilistic}.

Section \ref{sec:training} describes training, which estimates
parameters for the probabilistic database $P(w)$. This step first
automatically generates training data using the rule-based semantic
parser. This data is used to formulate a matrix factorization problem
that is optimized to estimate the database parameters.

% In our approach, only
% the probabilistic database $P(w)$ must be learned. Training data, in
% the form of logical forms with fully-observed entity arguments, is
% automatically generated by applying the open-vocabulary semantic
% parser to entity-linked text. The probabilistic database is
% parameterized as a matrix factorization model, and is trained by
% optimizing a ranking objective using stochastic gradient descent.

% Note that this probabilistic model is the exact opposite of the
% traditional semantic parsing model. Traditionally, $P(\ell | s)$ is
% learned, and $P(W)$ is a point distribution. Our system is exactly the
% opposite, using a \emph{deterministic} semantic parser and a
% \emph{learned} database.

\section{Rule-Based Semantic Parser}
\label{sec:rule}

The first part of our compositional semantics system is a rule-based
system that deterministically computes a logical form $\ell$ for a
text $s$. This component is used during inference to analyze the
logical structure of text, and during training to generate training
data (see Section \ref{sec:trainingdata}).  Several input/output pairs
for this system are shown in Figure \ref{fig:examplesem}.

The conversion to logical form has 3 phases:

\begin{enumerate}
\item \textbf{CCG syntactic parsing} parses the text and applies
  several deterministic syntactic transformations to facilitate
  semantic analysis.
\item \textbf{Entity linking} marks known Freebase entities in the
  text.
\item \textbf{Semantic analysis} assigns a logical form to each word,
  then composes them to produce a logical form for the complete text.
\end{enumerate}

\subsection{Syntactic Parsing}

The first step in our analysis is to syntactically parse the text. We
use the \textsc{ASP-SYN} parser \cite{krishnamurthy2014} trained on
CCGBank \cite{hockenmaier02ccgbank}. We then automatically transform
the resulting syntactic parse to make the syntactic structure more
amenable to semantic analysis. This step marks \ccgsyn{NP}s in
conjunctions by replacing their syntactic category with
\ccgsyn{NP[\mbox{conj}]}. This transformation allows semantic analysis to
distinguish between appositives and comma-separated lists. It also
transforms all verb arguments to core arguments, i.e., using the
category \ccgsyn{PP/NP} as opposed to \ccgsyn{((S\backslash
  NP)\backslash (S\backslash NP))/NP}. This step simplifies the
semantic analysis of verbs with prepositional phrase arguments. The
final transformation adds a word feature to each \ccgsyn{PP} category,
e.g., mapping \ccgsyn{PP} to \ccgsyn{PP[\mbox{by}]}. These features are used
to generate verb-preposition relation predicates, such as
$\pred{directed\_by}$.

\subsection{Entity Linking}

The second step is to identify mentions of Freebase entities in the
text. This step could be performed by an off-the-shelf entity linking
system \cite{ratinov2011,milne2008} or string matching. However, our
training and test data is derived from Clueweb 2009, so we rely on the
entity linking for this corpus provided by Gabrilovich et. al
\shortcite{gabrilovich2013}.

Our system incorporates the provided entity links into the syntactic
parse provided that they are consistent with the parse
structure. Specifically, we require that each mention is either (1) a
constituent in the parse tree with syntactic category $N$ or $NP$ or
(2) a collection of $N/N$ or $NP/NP$ modifiers with a single head
word. The first case covers noun and noun phrase mentions, while the
second case covers noun compounds. In both cases, we substitute a
single multi-word terminal into the parse tree spanning the mention
and invoke special semantic rules for mentions described in the next
section.

\subsection{Semantic analysis}

The final step uses the syntactic parse and entity links to produce a
logical form for the text. The system induces a logical form for every
word in the text based on its syntactic CCG category. Composing these
logical forms according to the syntactic parse produces a logical form
for the entire text.

\begin{figure}
{\small
\ul{Dan Hesse}, CEO of \ul{Sprint} \\
$\lambda x.\exists y.x = \pred{/en/dan\_hesse} \wedge \pred{ceo}(x) \wedge \pred{of}(x, y) \wedge y = \pred{/en/sprint}$ \\

\ul{Yankees} pitcher \\
$\lambda x.\exists y.\pred{pitcher}(x) \wedge \pred{nn}(y, x) \wedge y = \pred{/en/yankees}$ \\

\ul{Tom Cruise} plays \ul{Maverick} in the movie \ul{Top Gun}. \\
$\exists x,y,z. x = \pred{/en/tom\_cruise} \wedge \pred{plays}(x,
y) \wedge y = \pred{/en/maverick\_(character)} \wedge \pred{plays\_in}(x, z) \wedge z
= \pred{/en/top\_gun} $
}
\vspace{-.1in}
\caption{Example input/output pairs for our semantic analysis
  system. Mentions of Freebase entities in the text are indicated by
  underlines.}
\vspace{-.1in}
\label{fig:examplesem}
\end{figure}

Our semantic analyses are based on a relatively na\"ive
model-theoretic semantics. We focus on language whose semantics can be
represented with existentially-quantified conjunctions of unary and
binary predicates, ignoring, for example, temporal scope and
superlatives. Generally, our system models nouns and adjectives as
unary predicates, and verbs and prepositions as binary
predicates. Special multi-word predicates are generated for
verb-preposition combinations. Entity mentions are mapped to the
mentioned entity in the logical form. We also created special rules
for analyzing conjunctions, appositives, and relativizing
conjunctions. The complete list of rules used to produce these logical
forms is available
online.\footnote{\url{http://rtw.ml.cmu.edu/tacl2015_csf}}

We made several notable choices in designing this component. First,
multi-argument verbs are analyzed using pairwise relations, as in the
third example in Figure \ref{fig:examplesem}. This analysis allows us
to avoid reasoning about entity triples (quadruples, etc.), which are
challenging for the matrix factorization due to sparsity. Second,
noun-preposition combinations are analyzed as a category and relation,
as in the first example in Figure \ref{fig:examplesem}. We empirically
found that combining the noun and preposition in such instances
resulted in worse performance, as it dramatically increased the
sparsity of training instances for the combined relations.
% Furthermore, in many
% cases our analysis is more reasonable than the combined analysis; for
% example, ``Dan Hesse of Sprint'' is an acceptable statement given the
% first example in Figure \ref{fig:examplesem}.
Third, entity mentions with the $\ccgsyn{N/N}$ category are analyzed
using a special noun-noun relation, as in the second example in Figure
\ref{fig:examplesem}. Our intuition is that this relation shares
instances with other relations (e.g., ``city in Texas'' implies
``Texan city''). Finally, we lowercased each word to create its
predicate name, but performed no lemmatization or other normalization.

\subsection{Discussion}

The scope of our semantic analysis system is somewhat limited relative
to other similar systems \cite{bos2008,lewis2013} as it only outputs
existentially-quantified conjunctions of predicates. Our goal in
building this system was to analyze noun phrases and simple sentences,
for which this representation generally suffices. The reason for this
focus is twofold. First, this subset of language is sufficient to
capture much of the language surrounding Freebase entities. Second,
for various technical reasons, this restricted semantic representation
is easier to use (and more informative) for training the probabilistic
database (see Section \ref{sec:qro}).

% This semantic analysis system may appear somewhat na\"ive in
% comparison to other similar systems \cite{bos2008,lewis2013}, as it
% does not even attempt to model many linguistic phenomena (TODO:
% example). Our goal in building this system was to analyze noun phrases
% and simple sentences, for which it produces reasonable analyses.
% There are many reasons for this emphasis. First, these are the kinds
% of constructions in which Freebase entities appear. Second, it makes
% sense to evaluate a new approach to compositional semantics on
% (relatively) simple language. Third, complex logical forms are less
% useful for training. Fourth, the current version of our probabilistic
% database learning is restricted to categories and relations. TODO:
% existentially quantified conjunctions of predicates.

Note that this system can be straightforwardly extended to model
additional linguistic phenomena, such as additional logical operators
and generalized quantifiers, by writing additional rules. The
semantics of logical forms including these operations are well-defined
in our model, and the system does not even need to be re-trained to
incorporate these additions.

% We demonstrate this capability by adding additional rules for
% negation (TODO).

% TODO: Describe the scope / intended coverage of this system. Cite
% BOxer / lewis and steedman. Explicitly discuss quantifiers (e.g., 5
% movies). Formal logical restrictions and size of l.f.s (can reference
% later table).

% The
% semantics of more complex logical forms,  are well-defined by the probabilistic database
% 
% The
% focus of this paper, however, is on evaluating the combination of a
% compositional semantics
% 
% Prior work has clearly demonstrated that more sophisticated systems
% are possible.
% 
% Thus,  If 
% 
% Our goal in building this system is not to produce a complete,
% logically accurate semantic analysis of language, but rather a good
% enough analysis for the purpose of evaluating our compositional
% semantics approach.
% 
% This semantic analysis system is somewhat na\"ive
% 
% The focus of this semantic analysis system is on analyzing noun
% phrases and simple sentences. 
% 
% The output of this semantic analysis system is, typically, an
% existentially-quantified conjunction of predicates. Although this
% representation may seem fairly limited, it is sufficient for the types
% of language we expect to be able to analyze. 
% 
% Despite the open predicate vocabulary, the
% denotations of many words cannot be accurately represented as sets of
% Freebase entities.
% 
% Note that this system can be extended quite easily with additional
% rules for generalized quantifiers (e.g, ``5 movies''). To verify this,
% we annotated a rule for a simple kind of negation (TODO: demo?).
% 
% The design of this semantic analysis system is influenced by the kinds
% of language we expect to encounter 
% 
% Our goals in designing this semantic analysis system were 
% 
% 
%  A simple form of negation was included for the sole purpose
% of demonstrating that
% 
% The intended scope of this semantic analysis system is quite limited:
% it focuses on analyzing noun phrases and simple sentences with
% verbs. 
% 
% Generalized quantifiers
% 
% Although this semantic analysis system is not very sophisticated, it
% is sufficient for the purposes of this paper. Specifically, not all
% language can be semantically represented using
% 
% During training, it is much more efficient to 
% 
% Other systems for converting syntactic CCG parses to logical form are
% considerably more sophisticated than ours \cite{bos2008,lewis2013}.
%

% Note that we make no attempt to analyze negation or quantifiers
% (outside of conjunctions). These phenomena could be included in our
% analysis by annotating special logical forms for a small subset of
% words, such as ``not.'' We found that such quantifiers were
% empirically not very common in our evaluation, and thus we chose not
% to annotate them. Complexly quantified statements would also be
% omitted from the probabilistic database training, so there is little
% consequence in omitting these entries.

\section{Probabilistic Database}
\label{sec:probdb}

The second part of our compositional semantics system is a
probabilistic database. This database represents a distribution over
\emph{possible worlds}, where each world is an assignment of truth
values to every predicate instance. Equivalently, the probabilistic
database can be viewed as a distribution over databases or knowledge
bases.

    % In traditional semantic parsing, a
   % logical form can be evaluated against a world to produce a
  % denotation. Evaluating a logical form against a probabilistic database
 % produces a \emph{distribution} over denotations.

Formally, a probabilistic database is a collection of random
variables, each of which represents the truth value of a single
predicate instance. Given entities $e \in E$, categories $c \in C$,
and relations $r \in R$ the probabilistic database contains boolean
random variables $c(e)$ and $r(e_1, e_2)$ for each category and
relation instance, respectively. All of these random variables are
assumed to be independent. Let a world $w$ represent an assignment of
truth values to all of these random variables, where $c(e) = w_{c,e}$
and $r(e_1, e_2) = w_{r, e_1, e_2}$. By independence, the probability
of a world can be written as:
\begin{eqnarray*}
P(w) & = & \prod_{e \in E} \prod_{c \in C} P( c(e) = w_{c,e}) \times \\
 & & \prod_{e_1\in E} \prod_{e_2 \in E} \prod_{r \in R} P(r(e_1, e_2) = w_{r, e_1, e_2})
\end{eqnarray*}

The next section discusses how probabilistic matrix factorization is
used to model the probabilities of these predicate instances.

\subsection{Matrix Factorization Model}

The probabilistic matrix factorization model treats the truth of each
predicate instance as an independent boolean random variable that is
true with probability:
\begin{eqnarray*}
P( c(e) = \true{}) & = & \sigma ( \theta_c^T \phi_e ) \\
P( r(e_1,e_2) = \true{}) & = & \sigma ( \theta_r^T \phi_{(e_1,e_2)} )
\end{eqnarray*}

Above, $\sigma (x) = \frac{e^x}{1 + e^x}$ is the logistic function. In
these equations, $\theta_c$ and $\theta_r$ represent $k$-dimensional
vectors of per-predicate parameters, while $\phi_e$ and $\phi_{(e_1,
  e_2)}$ represent $k$-dimensional vector embeddings of each entity
and entity pair. This model contains a low-dimensional embedding of
each predicate and entity such that each predicate's denotation has a
high probability of containing entities with nearby vectors. The
probability that each variable is false is simply 1 minus the
probability that it is true.

This model can be viewed as matrix factorization, as depicted in
Figure \ref{fig:overview}. The category and relation instance
probabilities can be arranged in a pair of matrices of dimension $|E|
\times |C|$ and $|E|^2 \times |R|$. Each row of these matrices
represents an entity or entity pair, each column represents a category
or relation, and each value is between 0 and 1 and represents a truth
probability (Figure \ref{fig:overview}, bottom right). These two
matrices are factored into matrices of size $|E| \times k$ and $k
\times |C|$, and $|E|^2 \times k$ and $k \times |R|$, respectively,
containing $k$-dimensional embeddings of each entity, category, entity
pair and relation (Figure \ref{fig:overview}, bottom left). These
low-dimensional embeddings are represented by the parameters $\phi$
and $\theta$.

% In the first equation above, $\theta$ is a $|C| \times k$ matrix
% containing a $k$-dimensional vector embedding $\theta_c$ for each
% category $c$, and $\phi$ an $|E| \times k$ matrix containing a
% $k$-dimensional vector embedding $\phi_e$ for each entity $e$. The
% $\theta$ and $\phi$ matrices are defined analogously for relations and
% entity pairs in the second equation.

\section{Inference: Computing Marginal Probabilities}
\label{sec:inference}

Inference computes the marginal probability, over all possible worlds,
that each entity is an element of a text's denotation. In many cases
-- depending on the text -- these marginal probabilities can be
computed exactly in polynomial time.

The inference problem is to calculate $P(e \in \gamma | s)$ for each
entity $e$. Because both the semantic parser $P(\ell | s)$ and query
evaluation $P(\gamma | \ell, w)$ are deterministic, this problem can
be rewritten as:
\begin{eqnarray*}
P(e \in \gamma | s) & = & \sum_\gamma \textbf{1}(e \in \gamma) P(\gamma | s) \\
% & = & \sum_\gamma \sum_w \textbf{1}(e \in \gamma) P(\gamma | \ell, w) P(w)  \\
& = & \sum_w \textbf{1}(e \in \llbracket \ell \rrbracket_w) P(w) 
\end{eqnarray*}

Above, $\ell$ represents the logical form for the text $s$ produced by
the rule-based semantic parser, and \textbf{1} represents the
indicator function. The notation $\llbracket \ell \rrbracket_w$
represents denotation produced by (deterministically) evaluating the
logical form $\ell$ on world $w$. This inference problem corresponds
to query evaluation in a probabilistic database, which is \#P-hard in
general. Intuitively, this problem can be difficult because $P(\gamma
| s)$ is a joint distribution over sets of entities that can be
exponentially large in the number of entities.

However, a large subset of probabilistic database queries, known as
\emph{safe} queries, permit polynomial time evaluation
\cite{dalvi2007}. Safe queries can be evaluated extensionally using a
probabilistic notion of a denotation that treats each entity as
independent. Let $\pden{ \ell }$ denote a probabilistic denotation,
which is a function from entities (or entity pairs) to probabilities,
i.e., $\pden{ \ell }(e) \in [0, 1]$. The denotation of a
logical form is then computed recursively, in the same manner as a
non-probabilistic denotation, using probabilistic extensions of the
typical rules, such as:
\begin{eqnarray*}
\pden{c}(e) & = & \sum_w P(w) \textbf{1}(w_{c,e}) \\
\pden{r}(e_1, e_2) & = & \sum_w P(w) \textbf{1}(w_{r,e_1,e_2}) \\
\pden{c_1(x) \wedge c_2(x)}(e) & = & \pden{ c_1 }(e) \times \pden{ c_2 }(e) \\
\pden{\exists y.r(x, y)}(e) & = & 1 - \prod_{y \in E} (1 - \pden{r}(e, y)) \\
% P_{ext}(c_1(e_1) \wedge r(e_1, e_2)) & = & P(c(e_1)) \times P(r(e_1, e_2)) \\
% P_{ext}(\exists y.r(e, y)) & = & ( 1 - \prod_{y \in E} [1 -  P(\pred{of}(e, y)) \times P(\pred{company}(y))]) \\
\end{eqnarray*}

The first two rules are base cases that simply retrieve predicate
probabilities from the probabilistic database. The remaining rules
compute the probabilistic denotation of a logical form from the
denotations of its parts.\footnote{This listing of rules is partial as
  it does not include, e.g., negation or joins between one-argument
  and two-argument logical forms. However, the corresponding rules are
  easy to derive.} The formula for the probabilistic computation on
the right of each of these rules is a straightforward consequence of
the (assumed) independence of entities. For example, the last rule
computes the probability of an \pred{or} of a set of independent
random variables (indexed by $y$) using the identity $A \vee B = \neg
(\neg A \wedge \neg B)$. For safe queries, $\pden{ \ell }(e) = P(e \in
\gamma | s)$, that is, the probabilistic denotation computed according
to the above rules \emph{is equal to} the marginal probability
distribution. In practice, all of the queries in the experiments are
safe, because they contain only one query variable and do not contain
repeated predicates. For more information on query evaluation in
probabilistic databases, we refer the reader to Suciu et
al. \shortcite{suciu2011probabilistic}.

% To evaluate a safe query, one first creates a query
% plan that determines the intermediate sets of entities to create. For
% example, to evaluate, $\lambda x. \exists y. \pred{ceo}(x) \wedge
% \pred{of}(x, y) \wedge \pred{company}(y)$, we can first evaluate
% $\lambda x,y.\pred{of}(x, y) \wedge \pred{company}(y)$ to get a set of
% entity pairs. Next, we quantify out $y$ to get the set of entities for
% $\lambda x.\exists y.\pred{of}(x, y) \wedge
% \pred{company}(y)$. Finally, we join this with $\lambda
% x.\pred{ceo}(x)$ to get the final result.
% 
% The fundamental idea of safe query evaluation is to approximate the
% joint distributions over sets of entities with
% 
% 
% 
% Evaluating a probabilistic database query can be
% intractable because $P(\gamma | s)$ is a distribution over \emph{sets}
% of entities, which can be exponentially large in the number of
% entities. However, this distribution often factors such that the set
% membership of each entity is independent. Safe queries are those for
% which treating this distribution as if it factors results in the
% correct marginal probabilities.
% 
% \begin{figure}
% \begin{pspicture}(0,0)(7,5)
% \psframe(0.5,3.5)(1,4.5)
% \psframe(0.5,3.5)(1,4.5)
% \end{pspicture}
% \end{figure}
% 
% Safe queries can be evaluated similarly to standard database queries
% by maintaining additional probability tables for each intermediate
% result. Consider the logical form $\lambda x. \exists y. \pred{ceo}(x)
% \wedge \pred{of}(x, y) \wedge \pred{company}(y)$. The probabilistic
% database gives us a table of probabilities for each predicate,
% $P_{ext}(\pred{company}(e))$, etc.
% 
% for each predicate
% and intermediate result in the computation. , for each predicate or
% intermediate recursively by maintaining a list of probabilities
%

% TODO: marginal probabilities can also be approximated quite accurately for unsafe queries.

% These queries can be
% evaluated using simple recursive probability computations. For
% example, an entity $e$ is in the denotation of $\lambda
% x.\pred{ceo}(x) \wedge \pred{of}(x, \pred{/en/sprint})$ with
% probability $P(\pred{ceo}(e)) \times P(\pred{of}(e,
% \pred{/en/sprint}))$. Queries with existential (universal) quantifiers
% can also be handled by computing an \pred{or} (\pred{and}) over all
% possible entity values of the quantified variable. For example, an
% entity $e$ is in the denotation of $\lambda x. \exists
% y. \pred{ceo}(x) \wedge \pred{of}(x, y) \wedge \pred{company}(y)$ with
% probability $P(\pred{ceo}(e)) \times ( 1 - \prod_{y \in E} [1 -
%   P(\pred{of}(e, y)) \times P(\pred{company}(y))])$; this expression
% computes the \pred{or} over entities $y$ using the identity $A \vee B
% = \neg (\neg A \wedge \neg B)$.

Note that inference does not compute the most probable denotation,
$\max_\gamma P(\gamma | s)$. In some sense, the most probable
denotation is the correct output for a model-theoretic
semantics. However, it is highly sensitive to the probabilities in the
database, and in many cases it is empty (because a conjunction of
independent boolean random variables is unlikely to be
true). Producing a ranked list of entities is also useful for
evaluation purposes.

% 
% 
% \begin{eqnarray*}
% P(\gamma | s) = \sum_w \sum_\ell P(\gamma | \ell, w) P(w) P(\ell | s) 
% \end{eqnarray*}
% 
% 
% Consider query evaluation against a fixed world (database) $W$ as in
% traditional semantic parsing. A logical form $\ell$ can be
% deterministically evaluated against this world to return a denotation
% $\gamma$. This process is traditionally written as 
% 
% In the probabilistic case, there is no longer a single world, but
% rather a distribution $P(W)$ over possible worlds. However, logical
% forms can be evaluated against each possible world in the same fashion
% as above, producing a distribution over denotations:
% 
% \begin{eqnarray*}
% P(\gamma | \ell ; \theta, \phi) & = & \sum_W P( \gamma | \ell, W) P(W ; \theta, \phi) \\
%  & = & \sum_W \textbf{1}(\gamma = \llbracket \ell \rrbracket_W) P(W ; \theta, \phi) \\
% \end{eqnarray*}
% 
% Above, \textbf{1} represents the indicator function. The conditional
% probability distribution can be replaced with this indicator because
% evaluation is deterministic.
% 
% At first glance, the highest probability denotation appears to be the
% correct output for the model. However, this denotation is highly
% sensitive to the probabilities in the database, and in many cases is
% completely empty. Instead, for the evaluation, we compute the marginal
% probability that each entity is an element of the denotation:
% 
% \begin{eqnarray*}
% P(e \in \gamma | \ell; \theta, \phi) = \sum_\gamma \textbf{1}(e \in \gamma) P(\gamma | \ell ; \theta, \phi)
% \end{eqnarray*}

% This probability is a score for each entity that allows us to produce
% a ranked list of answers for each query. As it turns out, there is a
% considerable amount of work on computing these marginal probabilities
% in probabilistic databases; for more information, we refer the reader
% to Suciu et al. \shortcite{suciu2011probabilistic}.  Computing
% marginal probabilities for probabilistic database queries is \#P-hard
% in general, but many queries permit tractable evaluation
% \cite{dalvi2007}.
% 
% For example, an entity $e$ is in the denotation of $\lambda
% x.\pred{ceo}(x) \wedge \pred{of}(x, \pred{/en/sprint})$ with
% probability $P(\pred{ceo}(e)) \times P(\pred{of}(e,
% \pred{/en/sprint}))$. Queries with existential (universal) quantifiers
% can also be handled by computing an \pred{or} (\pred{and}) over all
% possible entity values of the quantified variable. For example, an
% entity $e$ is in the denotation of $\lambda x. \exists
% y. \pred{ceo}(x) \wedge \pred{of}(x, y) \wedge \pred{company}(y)$ with
% probability $P(\pred{ceo}(e)) \times ( 1 - \prod_{y \in E} [1 -
%   P(\pred{of}(e, y)) \times P(\pred{company}(y))])$; this expression
% computes the \pred{or} over entities $y$ using the identity $A \vee B
% = \neg (\neg A \wedge \neg B)$. For more information about query
% evaluation in probabilistic databases, we refer the reader to .

\section{Training}
\label{sec:training}

The training problem in our approach is to learn parameters $\theta$
and $\phi$ for the probabilistic database. We consider two different
objective functions for learning these parameters that use slightly
different forms of training data. In both cases, training has two
phases. First, we generate training data, in the form of observed
assertions or query-answer pairs, by applying the rule-based semantic
parser to a corpus of entity-linked web text. Second, we optimize the
parameters of the probabilistic database to rank observed assertions
or answers above unobserved assertions or answers.

% We work around the
% requirement for negative training data by using a novel ranking
% objective that trains the system to rank observed answers to queries
% above unobserved answers.

\subsection{Training Data}
\label{sec:trainingdata}

\begin{figure*}
\small

\textbf{Original sentence and logical form} \\
General \ul{Powell}, appearing Sunday on \ul{CNN} 's \ul{Late Edition}, said ... \\
{\small $\exists w,x,y,z. ~~~ w = \pred{/en/powell} \wedge \pred{general}(w) \wedge \pred{appearing}(w, x) \wedge \pred{sunday}(x) \wedge \pred{appearing\_on}(w, y) \wedge y = \pred{/en/late} \wedge \pred{'s}(z, y) \wedge z = \pred{/en/cnn} \wedge \pred{said}(w, ...)$ } \\

\textbf{Simplified logical form} \\
{\small $\exists w,y,z. ~~ w = \pred{/en/powell}  \wedge \pred{general}(w) \wedge \pred{appearing\_on}(w, y) \wedge y = \pred{/en/late} \wedge \pred{'s}(z, y) \wedge z = \pred{/en/cnn}$ } \\

{
\renewcommand{\tabcolsep}{0.1cm}
\begin{tabular}{l|ll}
\textbf{Instances} & \textbf{Queries} & \textbf{Answers} \\
{\small $\pred{general}(\pred{/en/powell})$ }& {\small $\lambda w.\pred{general}(w) \wedge \pred{appearing\_on}(w, \pred{/en/late})$ } & {\small $\pred{/en/powell}$ }\\
{\small $\pred{appearing\_on}(\pred{/en/powell}, \pred{/en/late})$ } & {\small $\lambda y.\pred{appearing\_on}(\pred{/en/powell}, y) \wedge \pred{'s}(\pred{/en/cnn}, y) $ } & {\small  $\pred{/en/late}$ }\\
{\small $\pred{'s}(\pred{/en/cnn}, \pred{/en/late})$ } & {\small $\lambda z.\pred{'s}(z, \pred{/en/late})$}& {\small $\pred{/en/cnn}$ }\\
\end{tabular}
}
\vspace{-.15in}
\caption{Illustration of training data generation applied to a single
  sentence. We generate two types of training data, predicate
  instances and queries with observed answers, by semantically parsing
  the sentence and extracting portions of the generated logical form
  with observed entity arguments. The predicate instances are
  extracted from the conjuncts in the simplified logical form, and the
  queries are created by removing a single entity from the simplified
  logical form.}
\label{fig:trainingdata}
\vspace{-.1in}
\end{figure*}

Training data is generated by applying the process illustrated in
Figure \ref{fig:trainingdata} to each sentence in an entity-linked web
corpus. First, we apply our rule-based semantic parser to the sentence
to produce a logical form. Next, we extract portions of this logical
form where every variable is bound to a particular Freebase entity,
resulting in a simplified logical form. Because the logical forms are
existentially-quantified conjunctions of predicates, this step simply
discards any conjuncts in the logical form containing a variable that
is not bound to a Freebase entity. From this simplified logical form,
we generate two types of training data: (1) predicate instances, and
(2) queries with known answers (see Figure \ref{fig:trainingdata}). In
both cases, the corpus consists entirely of assumed-to-be-true
statements, making obtaining negative examples a major challenge for
training.\footnote{A seemingly simple solution to this problem is to
  randomly generate negative examples; however, we empirically found
  that this approach performs considerably worse than both of the
  proposed ranking objectives.}

\subsection{Predicate Ranking Objective}
\label{sec:pro}

Riedel et al. \shortcite{riedel2013} introduced a ranking objective to
work around the lack of negative examples in a similar matrix
factorization problem. Their objective is a modified version of
Bayesian Personalized Ranking \cite{rendle2009} that aims to rank
observed predicate instances above unobserved instances.

This objective function uses observed predicate instances (Figure
\ref{fig:trainingdata}, bottom left) as training data. This data
consists of two collections, $\{(c_i, e_i)\}_{i=1}^n$ and $\{(r_j,
t_j)\}_{j=1}^m$, of observed category and relation instances. We use
$t_j$ to denote a tuple of entities, $t_j = (e_{j,1}, e_{j,2})$, to
simplify notation. The predicate ranking objective is:
\begin{eqnarray*}
O_P(\theta, \phi) & = & \sum_{i=1}^n \log \sigma( \theta_{c_i}^T (\phi_{e_i} - \phi_{e_i^\prime})) + \\
& & \sum_{j=1}^m \log \sigma( \theta_{r_j}^T (\phi_{t_j} - \phi_{t_j^\prime}))
\end{eqnarray*}

where $e_i^\prime$ is a randomly sampled entity such that $(c_i,
e_i^\prime)$ does not occur in the training data. Similarly,
$t_j^\prime$ is a random entity tuple such that $(r_j, t_j^\prime)$
does not occur. Maximizing this function attempts to find
$\theta_{c_i}$, $\phi_{e_i}$ and $\phi_{e_i^\prime}$ such that
$P(c_i(e_i))$ is larger than $P(c_i(e_i^\prime))$ (and similarly for
relations). During training, $e_i^\prime$ and $t_j^\prime$ are
resampled on each pass over the data set according to each entity or
tuple's empirical frequency.

\subsection{Query Ranking Objective}
\label{sec:qro}

The previous objective aims to rank the entities within each predicate
well. However, such within-predicate rankings are insufficient to
produce correct answers for queries containing multiple predicates --
the scores for each predicate must further be calibrated to work well
with each other given the independence assumptions of the
probabilistic database.

We introduce a new training objective that encourages good rankings
for entire queries instead of single predicates. The data for this
objective consists of tuples, $\{(\ell_i, e_i)\}_{i=1}^n$, of a query
$\ell_i$ with an observed answer $e_i$ (Figure \ref{fig:trainingdata},
bottom right). Each $\ell_i$ is a function with exactly one entity
argument, and $\ell_i(e)$ is a conjunction of predicate instances. For
example, the last query in Figure \ref{fig:trainingdata} is a function
of one argument $z$, and $\ell(e)$ is a single predicate instance,
$\pred{'s}(e, \pred{/en/late})$. The new objective aims to rank the
observed entity answer above unobserved entities for each query: {
% \setlength{\abovedisplayskip}{-5pt}
\setlength{\belowdisplayskip}{-10pt}
% \setlength{\abovedisplayshortskip}{0pt}
% \setlength{\belowdisplayshortskip}{0pt}
\begin{eqnarray*}
O_{Q}(\theta, \phi) & = & \sum_{i=1}^n \log P_{rank}(\ell_i, e_i, e_i^\prime)  \\
\end{eqnarray*}
}
% The expression $\theta_{c}^T (\phi_{e} - \phi_{e^\prime})$ in the
% predicate ranking objective can be viewed as an approximation of the
% probability that $e$ is ranked above $e^\prime$ in category $c$. Our
% goal is to extend this approximation to work for complete logical
% expressions. Our approach is to can define a new probabilistic
% database, $\pred{DB}_{e^\prime}$, where the random variable for each
% predicate instance $c(e)$ defines the (approximate) probability. Then the
% predicate ranking objective is simply:
% 
% $P_{\pred{DB}_{e^\prime}}(c(e))$
%% For example,
%% given the phrase ``Dan Hesse, CEO of Sprint'', we would produce the
%% logical form $\pred{ceo}(\pred{/en/dan\_hesse}) \wedge
%% \pred{of}(\pred{/en/dan\_hesse}, \pred{/en/sprint})$, and the training
%% example $(\ell = \lambda x.\pred{ceo}(x) \wedge \pred{of}(x,
%% \pred{/en/sprint}), e = \pred{/en/dan\_hesse})$.

$P_{rank}$ generalizes the approximate ranking probability defined by
the predicate ranking objective to more general queries. The
expression $\sigma (\theta_{c}^T (\phi_{e} - \phi_{e^\prime}))$ in the
predicate ranking objective can be viewed as an approximation of the
probability that $e$ is ranked above $e^\prime$ in category
$c$. $P_{rank}$ uses this approximation for each individual predicate
in the query. For example, given the query $\ell = \lambda x.c(x)
\wedge r(x, y)$ and entities $(e, e^\prime)$, $P_{rank}(\ell, e,
e^\prime) = \sigma( \theta_c (\phi_e - \phi_{e^\prime})) \times
\sigma( \theta_r (\phi_{(e,y)} - \phi_{(e^\prime, y)}))$. For this
objective, we sample $e^\prime_i$ such that $(\ell_i, e^\prime_i)$
does not occur in the training data.

When $\ell$'s body consists of a conjunction of predicates, the query
ranking objective simplifies considerably. In this case, $\ell$ can be
described as three sets of one-argument functions: categories
$C(\ell)= \{\lambda x.c(x)\}$, left arguments of relations $R_L(\ell)
= \{\lambda x.r(x, y)\}$, and right arguments of relations $R_R(\ell)
= \{\lambda x.r(y, x)\}$. Furthermore, $P_{rank}$ is a product so we
can distribute the log:
\begin{eqnarray*}
O_{Q}(\theta, \phi) & = & \sum_{i=1}^n \sum_{\lambda x.c(x) \in C(\ell_i)} \hspace{-.2in} \log
\sigma (\theta_{c} (\phi_{e_i} - \phi_{e^\prime_i})) \\
& & \hspace{-.2in} + \hspace{-.2in} \sum_{\lambda x.r(x, y) \in R_L(\ell_i)} \hspace{-.2in} \log \sigma (\theta_{r} ( \phi_{(e_i,y)} - \phi_{(e^\prime_i, y)} )) \\
& & \hspace{-.2in} + \hspace{-.2in} \sum_{\lambda x.r(y, x) \in R_R(\ell_i)} \hspace{-.2in} \log \sigma (\theta_{r} ( \phi_{(y, e_i)} - \phi_{(y, e^\prime_i)} )) 
\end{eqnarray*}

This simplification reveals that the main difference between $O_{Q}$
and $O_{P}$ is the sampling of the unobserved entities $e^\prime$ and
tuples $t^\prime$. $O_{P}$ samples them in an unconstrained fashion
from their empirical distributions for every predicate. $O_{Q}$
considers the larger context in which each predicate occurs, with two
major effects. First, more negative examples are generated for
categories because the logical forms $\ell$ are more specific. For
example, both ``president of Sprint'' and ``president of the US''
generate instances of the $\pred{president}$ predicate; $O_Q$ will use
entities that only occur with one of these as negative examples for
the other. Second, the relation parameters are trained to rank tuples
with a shared argument, as opposed to tuples in general.

Note that, although $P_{rank}$ generalizes to more complex logical
forms than existentially-quantified conjunctions, training with these
logical forms is more difficult because $P_{rank}$ is no longer a
product. In these cases, it becomes necessary to perform inference
within the gradient computation, which can be expensive. The
restriction to conjunctions makes inference trivial, enabling the
factorization above.

\section{Evaluation}
\label{sec:evaluation}

We evaluate our approach to compositional semantics on a question
answering task. Each test example is a (compositional) natural
language question whose answer is a set of Freebase entities. We
compare our open domain approach to several baselines based on prior
work, as well as a human-annotated Freebase query for each example.

\subsection{Data}

We used Clueweb09 web
corpus\footnote{\url{http://www.lemurproject.org/clueweb09.php}} with the
corresponding Google FACC entity linking \cite{gabrilovich2013} to
create the training and test data for our experiments. The training
data is derived from 3 million webpages, and contains 1.7m predicate
instances, 1.1m queries, 150k entities and 130k entity
pairs. Predicates that appeared fewer than 6 times in the training
data were replaced with the predicate \pred{UNK}, resulting in 16k
categories and 1.4k relations.

Our test data consists of fill-in-the-blank natural language questions
such as ``Incan emperor \underline{\quad{}}'' or ``Cunningham directed
Auchtre's second music video \underline{\quad{}}.'' These questions
were created by applying the training data generation process (Section
\ref{sec:trainingdata}) to a collection of held-out webpages. Each
natural language question has a corresponding logical form query
containing at least one category and relation.

We chose not to use existing data sets for semantic parsing into
Freebase as our goal is to model the semantics of language that cannot
necessarily be modelled using the Freebase schema. Existing data sets,
such as Free917 \cite{cai2013acl} and WebQuestions \cite{berant2013},
would not allow us to evaluate performance on this subset of
language. Consequently, we evaluate our system on a new data set with
unconstrained language. However, we do compare our approach against
manually-annotated Freebase queries on our new data set (Section
\ref{sec:spfb}).

All of the data for our experiments is available online at {\small
  \url{http://rtw.ml.cmu.edu/tacl2015_csf}}.

\begin{table}
\centering
{\small
\begin{tabular}{lc}  \toprule
\# of questions & 220 \\
\quad{} Avg. \# of predicates / query & 2.77 \\
\quad{} Avg. \# of categories / query & 1.75 \\
\quad{} Avg. \# of relations / query & 1.02 \\  
\quad{} Avg. \# of answers / query & 1.92 \\ \midrule % \quad{} mean \# of answers & 3.63 \\ % 422 correct answers overall.  
\# of questions with $\geq 1$ answer & \multirow{2}{*}{116} \\
(found by at least one system) & \\
% \quad{} Avg. \# of answers / query with $>1$ answer & 3.63 \\
\bottomrule
\end{tabular}
}
\vspace{-.1in}
\caption{Statistics of the test data set.}
\label{table:statistics}
\vspace{-.15in}
\end{table}

\subsection{Methodology}

Our evaluation methodology is inspired by information retrieval
evaluations \cite{manning2008}. Each system predicts a ranked list of
100 answers for each test question. We then pool the top 30 answers of
each system and manually judge their correctness. The correct answers
from the pool are then used to evaluate the precision and recall of
each system. In particular, we compute average precision (AP) for each
question and report the mean average precision (MAP) across all
questions. We also report a weighted version of MAP, where each
question's AP is weighted by its number of annotated correct
answers. Average precision is computed as $\frac{1}{m}
\sum_{k=1}^m\precision(k) \times \correct(k)$, where $\precision(k)$
is the precision at rank $k$, $\correct(k)$ is an indicator function
for whether the $k$th answer is correct, and $m$ is the number of
returned answers (at most 100).

Statistics of the annotated test set are shown in Table
\ref{table:statistics}. A consequence of our unconstrained data
generation approach is that some test questions are difficult to
answer: of the 220 queries, at least one system was able to produce a
correct answer for 116. The remaining questions are mostly
unanswerable because they reference rare entities unseen in the
training data.

\subsection{Models and Baselines}

\begin{table}
\centering
{\small
\begin{tabular}{lcc} \toprule
 & MAP & Weighted MAP \\ \midrule
\cb{} & 0.224 & 0.266 \\
\clb{} & 0.246 & 0.296 \\ \midrule
\uspr{} & 0.299 & 0.473 \\
\usqr{} & 0.309 & 0.492 \\ \midrule
\epr{} & 0.391 & 0.614 \\ 
\eqr{} & 0.406 & 0.645 \\ \midrule
Upper bound & 0.527 & 1.0 \\ \bottomrule
\end{tabular}
}
\vspace{-.08in}
\caption{Mean average precision for our question answering task. The
  difference in MAP between each pair of adjacent models is
  statistically significant ($p < .05$) via the sign test.}
\label{table:results}
\vspace{-.15in}
\end{table}

We implemented two baseline models based on existing techniques. The
\pred{CorpusLookup} baseline answers test questions by directly using
the predicate instances in the training data as its knowledge
base. For example, given the query $\lambda x.\pred{ceo}(x) \wedge
\pred{of}(x, \pred{/en/sprint})$, this model will return the set of
entities $e$ such that $\pred{ceo}(e)$ and $\pred{of}(e,
\pred{/en/sprint})$ both appear in the training data. All answers
found in this fashion are assigned probability 1.

The \pred{Clustering} baseline first clusters the predicates in the
training corpus, then answers questions using the clustered
predicates. The clustering aggregates predicates with similar
denotations, ideally identifying synonyms to smooth over sparsity in
the training data. Our approach is closely based on Lewis and Steedman
\shortcite{lewis2013}, though is also conceptually related to
approaches such as DIRT \cite{lin2001} and USP
\cite{poon2009unsupervised}. We use the Chinese Whispers clustering
algorithm \cite{biemann2006} and calculate the similarity between
predicates as the cosine similarity of their TF-IDF weighted entity
count vectors. The denotation of each cluster is the union of the
denotations of the clustered predicates, and each entity in the
denotation is assigned probability 1.

We also trained two probabilistic database models, \uspr{} and
\usqr{}, using the two objective functions described in Sections
\ref{sec:pro} and \ref{sec:qro}, respectively. We optimized both
objectives by performing 100 passes over the training data with
AdaGrad \cite{duchi2011} using an L2 regularization parameter of
$\lambda=10^{-4}$. The predicate and entity embeddings have 300
dimensions. These parameters were selected on the basis of preliminary
experiments with a small validation set.

Finally, we observed that \clb{} has high precision but low recall,
while both matrix factorization models have high recall with somewhat
lower precision. This observation suggested that an ensemble of
\pred{CorpusLookup} and \pred{Factorization} could outperform either
model individually. We created two ensembles, \epr{} and \eqr{}, by
calculating the probability of each predicate as a 50/50 mixture of
each model's predicted probability.

\subsection{Results}

Table \ref{table:results} shows the results of our MAP evaluation, and
Figure \ref{fig:pr-curve} shows a precision/recall curve for each
model. The MAP numbers are somewhat low because almost half of the
test questions have no correct answers and all models get an average
precision of 0 on these questions. The upper bound on MAP is the
fraction of questions with at least 1 correct answer. Note that the
models perform well on the answerable questions, as reflected by the
ratio of the achieved MAP to the upper bound. The weighted MAP metric
also corrects for these unanswerable questions, as they are assigned 0
weight in the weighted average.

\begin{figure}
\vspace{-.15in}
\center
\begin{pspicture}(-1.25,0.7)(7,6.5)

\rput[c](2.5,1){{\small Recall}}
\rput[c](-1.1,3.9){\rotateleft{{\small Precision}}}

% \psframe(3.7,6.3)(6.5,4.3)
\psline[linecolor=red](4.5,6)(3.8,6)
\rput[l](4.6,6){{\footnotesize \pred{Ens. ($O_Q$)}}}
\psline[linecolor=blue](4.5,5.7)(3.8,5.7)
\rput[l](4.6,5.7){{\footnotesize \pred{Ens. ($O_P$)}}}
\psline[linecolor=green](4.5,5.4)(3.8,5.4)
\rput[l](4.6,5.4){{\footnotesize \pred{Fact. ($O_Q$)}}}
\psline[linecolor=orange](4.5,5.1)(3.8,5.1)
\rput[l](4.6,5.1){{\footnotesize \pred{Fact. ($O_P$)}}}
\psline[linecolor=gray,linestyle=dashed,dash=5pt 3pt](4.5,4.8)(3.8,4.8)
\rput[l](4.6,4.8){{\footnotesize \pred{C.Lookup}}}
\psline[linecolor=black,linestyle=dashed,dash=5pt 3pt](4.5,4.5)(3.8,4.5)
\rput[l](4.6,4.5){{\footnotesize \pred{Clustering}}}

\psset{xunit=5\psunit,yunit=6\psunit}
\psaxes[dx=.2,Dx=.2,dy=.1,Dy=.1,Oy=0.3,labels=all]{-}(0,0.3)(0,0.3)(1,1.01)
\fileplot[plotstyle=dots,linecolor=red]{curvedata/map_answerable/ensemble_joint.txt}
\fileplot[plotstyle=line,linecolor=red]{curvedata/map_answerable/ensemble_joint.txt}
\fileplot[plotstyle=dots,dotstyle=+,linecolor=blue]{curvedata/map_answerable/ensemble_ranking.txt}
\fileplot[linecolor=blue]{curvedata/map_answerable/ensemble_ranking.txt}

\fileplot[plotstyle=dots,dotstyle=square*,linecolor=green]{curvedata/map_answerable/joint.txt}
\fileplot[linecolor=green]{curvedata/map_answerable/joint.txt}
\fileplot[plotstyle=dots,dotstyle=triangle*,linecolor=orange]{curvedata/map_answerable/ranking.txt}
\fileplot[linecolor=orange]{curvedata/map_answerable/ranking.txt}

\fileplot[plotstyle=dots,linecolor=gray,linestyle=dashed]{curvedata/map_answerable/corpus_lookup.txt}
\fileplot[linecolor=gray,linestyle=dashed]{curvedata/map_answerable/corpus_lookup.txt}
\fileplot[plotstyle=dots,dotstyle=+,linecolor=black,linestyle=dashed]{curvedata/map_answerable/clustering.txt}
\fileplot[linecolor=black,linestyle=dashed]{curvedata/map_answerable/clustering.txt}
\end{pspicture}

\vspace{-.2in}
\caption{Averaged 11-point precision/recall curves for the 116
  answerable test questions.}
\vspace{-.2in}
\label{fig:pr-curve}
\end{figure}

% TODO: example queries and answers. Make a figure

These results demonstrate several findings. First, we find that both
\pred{Factorization} models outperform the baselines in both MAP and
weighted MAP. The performance improvement seems to be most significant
in the high recall regime (right side of Figure
\ref{fig:pr-curve}). Second, we find that the query ranking objective
$O_Q$ improves performance over the predicate ranking objective $O_P$
by 2-4\% on the answerable queries. The precision/recall curves show
that this improvement is concentrated in the low recall
regime. Finally, the ensemble models are considerably better than
their component models; however, even in the ensembled models, we find
that $O_Q$ outperforms $O_P$ by a few percent.

%% The results in Table \ref{table:results} also
%% show that both ensembles dramatically outperform their constituent
%% models. In both the \pred{Factorization} and \pred{Ensemble} models,
%% the query ranking objective improves performance over the predicate
%% ranking objective.

%%  presents an 11-point precision/recall curve
%% for each model. These curves show that both \pred{Factorization}
%% approaches uniformly outperform \clb{} and \cb{}, but the most
%% dramatic difference is in the high recall regime. We also see that the
%% query ranking objective $O_Q$ provides a consistent performance
%% improvement over the predicate ranking objective $O_P$.

\subsection{Comparison to Semantic Parsing to Freebase}
\label{sec:spfb}

A natural question is whether our open vocabulary approach outperforms
a closed approach for the same problem, such as semantic parsing to
Freebase (e.g., Reddy et al. \shortcite{reddy2014}). In order to
answer this question, we compared our best performing model to a
manually-annotated Freebase query for each test question. This
comparison allows us to understand the relative advantages of open and
closed predicate vocabularies.

The first author manually annotated a Freebase MQL query for each
natural language question in the test data set. This annotation is
somewhat subjective, as many of the questions can only be inexactly
mapped on to the Freebase schema. We used the following guidelines in
performing the mapping: (1) all relations in the text must be mapped
to one or more Freebase relations, (2) all entities mentioned in the
text must be included in the query, (3) adjective modifiers can be
ignored and (4) entities not mentioned in the text may be included in
the query. The fourth condition is necessary because many one-place
predicates, such as $\pred{mayor}(x)$, are represented in Freebase
using a binary relation to a particular entity, such as
$\pred{government\_office/title}(x, \pred{/en/mayor})$.

\begin{table}
\centering
{\small
\begin{tabular}{lc}  \toprule
\# of questions w/ an annotated MQL query & 142 \\
\quad{} query returns $>1$ answer & 95 \\
\quad{} query returns no answers & 47 \\  \midrule
\# of questions w/o an MQL query & 78 \\ \bottomrule
\end{tabular}
}
\vspace{-.1in}
\caption{Statistics of the Freebase MQL queries annotated for the test data set.}
\label{table:fbstatistics}
\vspace{-.1in}
\end{table}

\begin{table}
\centering
\vspace{-.04in}
{\small
\begin{tabular}{lcc} \toprule
 & MAP  \\ \midrule
\eqr{} & 0.263  \\
Freebase & 0.385 \\ \bottomrule
\end{tabular}
}
\vspace{-.1in}
\caption{Mean average precision of our best performing model compared
  to a manually annotated Freebase query for each test question.}
\label{table:fbresults}
\vspace{-.1in}
\end{table}

Statistics of the annotated queries are shown in Table
\ref{table:fbstatistics}. Coverage is reasonably high: we were able to
annotate a Freebase query for 142 questions (65\% of the test set). The
remaining unannotatable questions are due to missing predicates in
Freebase, such as a relation defining the emperor of the Incan
empire. Of the 142 annotated Freebase queries, 95 of them return at
least one entity answer. The queries with no answers typically
reference uncommon entities which have few or no known relation
instances in Freebase. The annotated queries contain an average of
2.62 Freebase predicates.

We compared our best performing model, \eqr{}, to the manually
annotated Freebase queries using the same pooled evaluation
methodology. The set of correct answers contains the correct
predictions of \eqr{} from the previous evaluation along with all
answers from Freebase. 
% (That is, we assume that Freebase has perfect precision.)

Results from this evaluation are shown in Table
\ref{table:fbresults}.\footnote{The numbers in this table are not
  comparable to the numbers in Table \ref{table:results} as the
  correct answers for each question are different.} In terms of
overall MAP, Freebase outperforms our approach by a fair
margin. However, this initial impression belies a more complex
reality, which is shown in Table \ref{table:fbresults2}. This table
compares both approaches by their relative performance on each test
question. On approximately one-third of the questions, Freebase has a
higher AP than our approach. On another third, our approach has a
higher AP than Freebase. On the final third, both approaches perform
equally well -- these are typically questions where neither approach
returns any correct answers (67 of the 75). Freebase outperforms in
the overall MAP evaluation because it tends to return more correct
answers to each question.

Note that the annotated Freebase queries have several advantages in
this evaluation. First, Freebase contains significantly more predicate
instances than our training data, which allows it to produce more
complete answers. Second, the Freebase queries correspond to the
performance of a perfect semantic parser, while current semantic
parsers achieve accuracies around 68\% \cite{berant2014}.

The results from this experiment suggest that closed and open
predicate vocabularies are complementary. Freebase produces high
quality answers when it covers a question. However, many of the
remaining questions can be answered correctly using an open vocabulary
approach like ours. This evaluation also suggests that recall is a
limiting factor of our approach; in the future, recall can be improved
by using a larger corpus or including Freebase instances during
training.

\begin{table}
\centering
{\small
\begin{tabular}{lcc} \toprule
 & \# of queries  \\ \midrule
Freebase higher AP & 75 (34\%)  \\
equal AP & 75 (34\%) \\
\eqr{} higher AP & 70 (31\%) \\ \bottomrule
\end{tabular}
}
\vspace{-.1in}
\caption{Question-by-question comparison of model performance. Each
  test question is placed into one of the three buckets above,
  depending on whether Freebase or \eqr{} achieves a better average
  precision (AP) for the question.}
\label{table:fbresults2}
\vspace{-.1in}
\end{table}

% \subsection{Examples}
% TODO

\section{Related Work}
\label{sec:priorwork}

\subsubsection*{Open Predicate Vocabularies}

There has been considerable work on generating semantic
representations with an open predicate vocabulary. Much of the work is
non-compositional, focusing on identifying similar predicates and
entities. DIRT \cite{lin2001}, Resolver \cite{yates07resolver} and
other systems \cite{yao2012} cluster synonymous expressions in a
corpus of relation triples. Matrix factorization is an alternative
approach to clustering that has been used for relation extraction
\cite{riedel2013,yao2013} and finding analogies
\cite{turney2008,speer2008}. All of this work is closely related to
distributional semantics, which uses distributional information to
identify semantically similar words and phrases
\cite{turney2010,griffiths2007}.

Some work has considered the problem of compositional semantics with
an open predicate vocabulary. Unsupervised semantic parsing
\cite{poon2009unsupervised,titov2011} is a clustering-based approach
that incorporates composition using a generative model for each
sentence that factors according to its parse tree. Lewis and Steedman
\shortcite{lewis2013} also present a clustering-based approach that
uses CCG to perform semantic composition. This approach is similar to
ours, except that we use matrix factorization and Freebase entities.

Finally, some work has focused on the problem of textual inference
within this paradigm. Fader et al. \shortcite{fader2013} present a
question answering system that learns to paraphrase a question so that
it can be answered using a corpus of Open IE triples \cite{fader2011}.
Distributional similarity has also been used to learn weighted logical
inference rules that can be used for recognizing textual entailment or
identifying semantically similar text
\cite{garrette2011,garrette2013,beltagy2013}. This line of work
focuses on performing inference between texts, whereas our work
computes a text's denotation.

% Our  These works combine natural language and
% knowledge base predicates in a single matrix factorization to predict
% new instances of the knowledge base predicates. However, they do not
% define a compositional semantics for longer natural language phrases,
% as we do in our work. We also introduce a novel objective function
% that improves performance on compositional queries. Note that our work
% does not use a universal schema in the sense that it does not combine
% textual predicates and Freebase predicates in the same schema.

% Several systems have attempted to automatically generalize semantic
% representations using an open predicate vocabulary. However, existing
% approaches such as DIRT, USP, Resolver, Lewis and Steedman (TODO:
% cite) are based on clustering synonyms. Our dimensionality reduction
% approach is more flexible than clustering, as the learned embedding of
% predicates can represent (e.g.,) near-synonyms or synonyms in a
% particular context. Furthermore, the expressive power of CCG enables
% our approach to represent negation and other linguistic phenomena that
% cannot be represented in these existing systems (barring Lewis and
% Steedman).

A significant difference between our work and most of the related work
above is that our work computes denotations containing Freebase
entities. Using these entities has two advantages: (1) it enables us
to use entity linking to disambiguate textual mentions, and (2) it
facilitates a comparison against alternative approaches that rely on a
closed predicate vocabulary. Disambiguating textual mentions is a
major challenge for previous approaches, so an entity-linked corpus is
a much cleaner source of data. However, our approach could also work
with automatically constructed entities, for example, created by
clustering mentions in an unsupervised fashion \cite{singh11}.

% There is an interesting connection between this work and compositional
% semantics with vector space models
% \cite{socher2011,hermann2013,krishnamurthy2013vssp}. In both cases,
% the meanings of individual words are represented as vectors. However,
% vector space models perform semantic composition directly in the space
% of vectors (or matrices, etc.), while our approach first maps these
% vectors to sets of entities, then performs composition using logical
% operations on these sets. Our approach has the advantage of being able
% to straightforwardly model phenomena that are challenging for vector
% space models, such as the action of subordinating conjunctions, e.g.,
% ``that'' in ``car that I drive.''

\subsubsection*{Semantic Parsing}

Several semantic parsers have been developed for Freebase
\cite{cai2013acl,kwiatkowski2013,berant2013,berant2014}. Our approach
is most similar to that of Reddy et al. \shortcite{reddy2014}, which
uses fixed syntactic parses of unlabeled text to train a Freebase
semantic parser. Like our approach, this system
automatically-generates query/answer pairs for training. However, this
system, like all Freebase semantic parsers, uses a closed predicate
vocabulary consisting of only Freebase predicates. In contrast, our
approach uses an open predicate vocabulary and can learn denotations
for words whose semantics cannot be represented using Freebase
predicates. Consequently, our approach can answer many questions that
these Freebase semantic parsers cannot (see Section \ref{sec:spfb}).

The rule-based semantic parser used in this paper is very similar to
several other rule-based systems that produce logical forms from
syntactic CCG parses \cite{bos2008,lewis2013}. We developed our own
system in order to have control over the particulars of the analysis;
however, our approach is compatible with these systems as well.

\subsubsection*{Probabilistic Databases}

Our system assigns a model-theoretic semantics to statements in
natural language \cite{dowty81} using a learned distribution over
possible worlds. This distribution is concisely represented in a
probabilistic database, which can be viewed as a simple Markov Logic
Network \cite{richardson2006} where all of the random variables are
independent. This independence simplifies query evaluation:
probabilistic databases permit efficient exact inference for safe
queries \cite{suciu2011probabilistic}, and approximate inference for
the remainder \cite{gatterbauer2010,gatterbauer2015}.

\section{Discussion}
\label{sec:discussion}

This paper presents an approach for compositional semantics with an
open predicate vocabulary. Our approach defines a probabilistic model
over denotations (sets of Freebase entities) conditioned on an input
text. The model has two components: a rule-based semantic parser that
produces a logical form for the text, and a probabilistic database
that defines a distribution over denotations for each predicate. A
training phase learns the probabilistic database by applying
probabilistic matrix factorization with a query/answer ranking
objective to logical forms derived from a large, entity-linked web
corpus. An experimental analysis demonstrates that this approach
outperforms several baselines and can answer many questions that
cannot be answered by semantic parsing into Freebase.

Our approach learns a model-theoretic semantics for natural language
text tied to Freebase, as do some semantic parsers, except with an
open predicate vocabulary. This difference influences several other
aspects of the system's design. First, because no knowledge base with
the necessary knowledge exists, the system is forced to \emph{learn}
its knowledge base (in the form of a probabilistic database). Second,
the system can directly map syntactic CCG parses to logical forms, as
it is no longer necessary to map words to a closed vocabulary of
knowledge base predicates. In some sense, our approach is the exact
opposite of the typical semantic parsing approach: usually, the
semantic parser is learned and the knowledge base is fixed; here, the
knowledge base is learned and the semantic parser is fixed. From a
machine learning perspective, training a probabilistic database via
matrix factorization is easier than training a semantic parser, as
there are no difficult inference problems. However, it remains to be
seen whether a learned knowledge base can achieve similar recall as a
fixed knowledge base on the subset of language it covers.

There are two limitations of this work. The most obvious limitation is
the restriction to existentially quantified conjunctions of
predicates. This limitation is not inherent to the approach, however,
and can be removed in future work by using a system like Boxer
\cite{bos2008} for semantic parsing. A more serious limitation is the
restriction to one- and two-argument predicates, which prevents our
system from representing events and $n$-ary relations. Conceptually, a
similar matrix factorization approach could be used to learn
embeddings for $n$-ary entity tuples; however, in practice, the
sparsity of these tuples makes learning challenging. Developing
methods for learning $n$-ary relations is an important problem for
future work.

% In our evaluation, we noticed two common sources of error. The first
% is noisy training data due to entity linking and syntactic parsing
% errors. For example, processing a comma-separated list often results
% in noun-noun relations between its elements. The consequences of these
% errors are noticeable in the output of test questions as well. An
% improved semantic parsing component or a denoising step could improve
% performance. The second is that the matrix factorization tends to
% assign more extreme probabilities to more frequent entities as their
% vector embeddings have greater norms. This can occasionally produce
% oddly-ranked results. Normalizing the embeddings in some fashion may
% help.

% TODO: discussion of coverage / expressivity of the approach. Which
% language phenomena are covered by the representation? What kinds of
% negation?

A direction for future work is scaling up the size of the training
corpus to improve recall. Low recall is the main limitation of our
current system as demonstrated by the experimental analysis. Both
stages of training, the data generation and matrix factorization, can
be parallelized using a cluster. All of the relation instances in
Freebase can also be added to the training corpus. It should be
feasible to increase the quantity of training data by a factor of
10-100, for example, to train on all of ClueWeb. Scaling up the
training data may allow a semantic parser with an open predicate
vocabulary to outperform comparable closed vocabulary systems.

% An important open problem is extending the matrix factorization
% approach used here to learn denotations for higher-arity (e.g.,
% ternary) predicates. Such predicates are necessary to represent
% events. This problem is challenging due to sparsity in the
% co-occurrences of entity triples (quadruples, etc.), and because many
% event mentions will only reference a subset of the event's
% arguments. One possible approach is to introduce latent event
% variables and binary relations between events and their arguments.

% Another conclusion from this work is that there are complementary
% roles for both open and closed predicate vocabularies. An elegant way
% to combine these approaches would be to construct a semantic parser
% that is capable of using both Freebase predicates as well as
% open-domain predicates. This semantic parser would have higher
% coverage than current approaches, while still obtaining the advantages
% of Freebase. Another advantage is the ability to learn the semantic
% analysis component, which may improve the semantic analyses for tricky
% phenomena like quantifier scoping. These integrated methods are a
% promising direction for future work.

\section*{Acknowledgments}
This research was supported in part by DARPA under contract number
FA8750-13-2-0005, and by a generous grant from Google. We additionally
thank Matt Gardner, Ndapa Nakashole, Amos Azaria and the anonymous
reviewers for their helpful comments.

\bibliography{semantic_parsing}{}
\bibliographystyle{acl2012}

\end{document}