### 2016-03-21 papers/acl2016/paper.tex
Fix text width and remove commented-out text
Former-commit-id: 2ea804b60d3534a40b77dc2478323c370288cca5

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{times}
\usepackage{latexsym}
\usepackage{color}
\usepackage{booktabs}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\mattnote}[1]{\textcolor{red}{NOTE: #1}}
\newcommand{\blank}{\underline{\hspace{.5cm}}}
\newcommand{\lexicalpredicate}[1]{\ensuremath{\textit{#1}}}
\newcommand{\formalpredicate}[1]{\ensuremath{\textsc{#1}}}
\newcommand{\entity}[1]{\ensuremath{\textsc{#1}}}

\newcommand{\true}[0]{\formalpredicate{true}}
\newcommand{\false}[0]{\formalpredicate{false}}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

\title{Open-Vocabulary Semantic Parsing\\with both Distributional
Statistics and Formal Knowledge}

\author{}%Matt Gardner and Jayant Krishnamurthy\\
%Allen Institute for Artificial Intelligence\\
%Seattle, Washington, USA\\
%{\tt \{mattg,jayantk\}@allenai.org}}

\date{}

\begin{document}

\maketitle

\begin{abstract}

  We consider the problem of mapping language onto meaning representations that
  are executable, returning truth values for complete statements and
  distributions over entities for queries.  The typical approach to solve this
  problem is semantic parsing; however, this approach fails when the language
  cannot be represented by a single query over the Freebase schema.  Instead,
  our approach represents language as a \emph{weighted combination} of queries
  over the formal schema, along with a distributional component similar to a
  word embedding.  This technique combines the advantages of traditional
  semantic parsing (compositionality and the ability to leverage a structured
  knowledge base) with the broad coverage obtained by distributional methods.
  We present relative gains of over 120\% in mean average precision versus
  prior work on an open-vocabulary fill-in-the-blank question answering task.

\end{abstract}

\section{Introduction}

How should a computer program represent the meaning of a word or a phrase?
Recent work has suggested two vastly different approaches---distributional
vector space models and semantic parsing---each with its own advantages and
disadvantages. On one hand, distributional models produce a vector
representation of the meaning of each word from a large unlabeled text corpus.
These models have broad coverage, but only a shallow semantics centered around
word similarity. On the other hand, semantic parsing translates natural
language into logical queries against a formal knowledge base, such as
Freebase. This approach produces a rich compositional semantics and can
leverage structured information from the knowledge base. However, semantic
parsers have limited coverage as they can only represent the meanings of words
that can be mapped to knowledge base predicates. A critical question is
\emph{how do we obtain the benefits of both of these approaches?}

\begin{figure*}[ht]
  \small
  \begin{minipage}{0.18\linewidth}
    \textbf{Input Text}\\Italian architect \blank{}
  \end{minipage}
%
  ~~ $\longrightarrow$ ~~~~~~
%
  \begin{minipage}{0.7\linewidth}
    \textbf{Logical Form}\\
    $\lambda x.\lexicalpredicate{architect}(x)~\land~\lexicalpredicate{architect\_N/N}(\entity{Italy}, x)$
  \end{minipage}

  \vspace{.2in}
  \begin{minipage}{0.5\linewidth}
    \textbf{Category models}\\
    \begin{tabular}{@{}lll}
      \lexicalpredicate{architect}: & $\theta$: &[.2, -.6, \ldots] \\
      & $\omega$: & \textsc{type:architect} $\rightarrow$ .52 \\
      &           & \textsc{type:designer} $\rightarrow$ .32 \\
      &           & \textsc{nationality:Italy} $\rightarrow$ .20 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \textbf{Relation models}\\
    \begin{tabular}{@{}lll}
      \lexicalpredicate{architect\_N/N}: & $\theta$: &[-.9, .1, \ldots] \\
      & $\omega$: & \textsc{/person/nationality\textsuperscript{-1}} $\rightarrow$ .29 \\
      &           & \textsc{/structure/architect} $\rightarrow$ .11 \\
      &           & \textsc{/person/ethnicity\textsuperscript{-1}} $\rightarrow$ .05 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}

  \begin{minipage}{0.5\linewidth}
    \textbf{Entity models}\\
    \begin{tabular}{@{}lll}
      \entity{Palladio}: & $\phi$: &[.15, -.8, \ldots] \\
      & $\psi$: & \textsc{type:architect} $\rightarrow$ 1 \\
      &           & \textsc{nationality:Italy} $\rightarrow$ 1 \\
      &           & $\cdots$ \\
      \entity{Obama}: & $\phi$: &[.85, .1, \ldots] \\
      & $\psi$: & \textsc{type:politician} $\rightarrow$ 1 \\
      &           & \textsc{nationality:Italy} $\rightarrow$ 0 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \textbf{Entity pair models}\\
    \begin{tabular}{@{}lll}
      \entity{(Italy, Palladio)}: & $\phi$: &[-.8, .2, \ldots] \\
      & $\psi$: & \textsc{/person/nationality\textsuperscript{-1}} $\rightarrow$ 1 \\
      &           & \textsc{/structure/architect} $\rightarrow$ 0 \\
      &           & $\cdots$ \\
      \entity{(Italy, Obama)}: & $\phi$: &[-.2, -.2, \ldots] \\
      & $\psi$: & \textsc{/person/nationality\textsuperscript{-1}} $\rightarrow$ 0 \\
      &           & \textsc{/structure/architect} $\rightarrow$ 0 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}

  \begin{minipage}{0.5\linewidth}
    \textbf{Output}\\
    \begin{tabular}{@{}ll}
      $P(\entity{Palladio})$: & 0.79 \\
      $P(\entity{Obama})$: & 0.01 \\
      $\cdots$ \\
    \end{tabular}
  \end{minipage}

  \vspace{-.1in}
  \caption{Overview of the components of our model.  Given an input text, we
  use a CCG parser and an entity linker to produce a logical form that is a
  conjunction of predicates involving the entities in the text.  For each
  predicate (both categories and relations), we learn a distributional vector
  $\theta$, as well as weights $\omega$ associated with formal features from
  Freebase (which are shortened to fit in the figure).  For each entity and
  entity pair, we also learn a distributional vector $\phi$, and we extract a
  feature vector $\psi$ from Freebase.  These models are combined to assign a
  probability to candidate entities, assigning a high probability to
  \entity{Andrea Palladio} and a low probability to \entity{Barack Obama}.}
  \label{fig:overview}
  \vspace{-.1in}
\end{figure*}

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} provide a
partial answer to this question with their open-vocabulary semantic parser.
This system answers compositional, fill-in-the-blank natural language queries
against Freebase, such as ``Hollywood honcho \blank{},'' using
\emph{distributional} models of predicate semantics. The method semantically
parses text to a surface logical form containing predicates derived from the
text itself, such as $\lambda x.\lexicalpredicate{honcho}(x) \land
\lexicalpredicate{honcho\_N/N}(\entity{Hollywood}, x)$, then learns the
denotations of these predicates using distributional information from a large,
entity-linked corpus. This method combines the broad coverage of distributional
approaches with the compositionality of semantic parsing; however, it does not
use structured information from Freebase to assist with its predictions.

In this work, we introduce new models of predicate semantics that leverage both
structured information from Freebase and distributional information from a
corpus. Our key observation is that, while there is no Freebase predicate that
corresponds exactly to the meaning of ``honcho,'' Freebase \emph{does} contain
predicates that \emph{partially} represent its meaning. For example, an entity
that can be called ``honcho'' is likely to have type \formalpredicate{person}.
We therefore incorporate \emph{features} based on Freebase paths into the
predicate models of Krishnamurthy and Mitchell and learn per-predicate weights
for these features using distributional information (\figref{overview}). This
approach effectively learns the meaning of a word as a distributional vector
plus a \emph{weighted combination of Freebase queries}, a considerably more
expressive representation than either traditional semantic parsers or the work
by Krishnamurthy and Mitchell.

We demonstrate our approach on the task of answering fill-in-the-blank natural
language queries.  We first present two simple improvements to the
distributional models of Krishnamurthy and Mitchell (an improved surface
logical form representation, and better candidate generation during inference),
improving the performance of that system by over 40\%.  We then show that our
approach improves mean average precision by 62\% over this improved baseline.
Taken together, the methods presented in this paper give a relative gain over
prior work of more than 120\%.

\section{Background}
\label{sec:background}

In this work, we make heavy use of the open-vocabulary semantic parser of
Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}, as well as
features derived from a knowledge base using a technique called subgraph
feature extraction~\cite{gardner-2015-sfe}.  We briefly describe both of these
techniques.

\subsection{Open-vocabulary semantic parsing}
\label{sec:jayant-semparse}

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} introduced a
technique for learning semantic parsers with open predicate vocabularies.
Instead of mapping text to Freebase queries, this method parses text to a
surface logical form whose predicates are derived directly from the words in
the text (an example of these logical forms is shown in \figref{overview}).
Next, a distribution over denotations for each predicate is learned using a
matrix factorization approach similar to that of Riedel et
al.~\shortcite{riedel-2013-mf-universal-schema}. This distribution is concisely
represented using a probabilistic database, which also enables efficient
probabilistic execution of logical form queries.

The truth probability of a predicate instance such as
\lexicalpredicate{architect}(\entity{AndreaPalladio}) is learned using
probabilistic matrix factorization. This factorization has two sets of
parameters: each category or relation has a learned $k$-dimensional embedding
$\theta$, and each entity or entity pair has a learned $k$-dimensional
embedding $\phi$. The probability assigned to a category instance $c(e)$ or
relation instance $r(e_1, e_2)$ is given by:

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e ) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} )
\end{align*}

The probability of a predicate instance is the (sigmoided) inner product of the
corresponding predicate and entity embeddings. Thus, predicates with nearby
embeddings will have similar distributions over the entities in their
denotation. The parameters $\theta$ and $\phi$ are learned using a query
ranking objective that optimizes them to rank entities observed in the
denotation of a logical form above unobserved entities. Note that this approach
is similar to modern techniques for constructing word embeddings in that it
factors a cooccurrence matrix between predicates and
entities~\cite{levy-2014-w2v-as-mf}. Given the trained predicate and entity
parameters, the system is capable of efficiently computing the marginal
probability that an entity is an element of a logical form's denotation using
approximate inference algorithms for probabilistic databases.

In this paper, we use the open-vocabulary semantic parsing framework outlined
in this section.  However, where Krishnamurthy and Mitchell used only
distributional methods to compute $P(c(e))$ and $P(r(e_1, e_2))$, we augment
these models with features derived from a formal knowledge base, as described
in \secref{method}.

\subsection{Subgraph feature extraction}

\begin{figure}
  {\center
  \begin{tikzpicture}[
    ->,
    shorten >=1pt,
    auto,
    node distance=3cm,
    thick,
    main node/.style={draw=none,fill=none}
    ]

    \node[main node] (1) {\entity{Palladio}};
    \node[main node] (2) [right=4cm of 1] {\entity{Italy}};
    \node[main node] (3) [below=1.7cm of 1] {\entity{architect}};
    \node[main node] (4) [below=1.7cm of 2] {\entity{country}};
    \node[main node] (5) [below right=.1cm and .5cm of 3] {\entity{Villa Capra}};

    \path[]
      (1) edge node [sloped, anchor=center, above] {\formalpredicate{nationality}} (2)
      (1) edge node [sloped, anchor=center, above] {\formalpredicate{type}} (3)
      (1) edge node [sloped, anchor=center, above] {\formalpredicate{designed}} (5)
      (2) edge node [sloped, anchor=center, above] {\formalpredicate{type}} (4)
      (5) edge node [sloped, anchor=center, above] {\formalpredicate{located\_in}} (2);
  \end{tikzpicture}
  }
  \textbf{Features between \entity{Palladio} and \entity{Italy}}:\\
  \textless\formalpredicate{nationality}\textgreater\\
  \textless\formalpredicate{designed}$\rightarrow$\formalpredicate{located\_in}\textgreater

  \textbf{Features for \entity{Palladio}}:\\
  \textless\formalpredicate{nationality}\textgreater\\
  \textless\formalpredicate{nationality}\textgreater:\entity{Italy}\\
  \textless\formalpredicate{type}\textgreater:\entity{Architect}\\
  \textless\formalpredicate{designed}$\rightarrow$\formalpredicate{located\_in}\textgreater \\
  \textless\formalpredicate{designed}$\rightarrow$\formalpredicate{located\_in}\textgreater:\entity{Italy}

  \caption{A subset of the Freebase graph, and some example extracted features.
  The actual Freebase relations and entity identifiers used are modified here
  to aid readability.}
  \label{fig:sfe}
\end{figure}

Subgraph feature extraction (SFE) is a technique introduced by Gardner and
Mitchell~\shortcite{gardner-2015-sfe} for generating feature matrices over node
pairs in a graph with labeled edges.  Given a pair of nodes in a graph, SFE
performs a search to characterize and extract features from a subgraph around
those two nodes.  Example features include the set of edge sequences that
connect the two nodes.  \figref{sfe} shows an example subgraph for the pair
\entity{Andrea Pallado} and \entity{Italy}, and features that would be
extracted for that entity pair.  Gardner and Mitchell used these features to
perform link prediction in the graph, using characteristics of the subgraph
around two nodes to determine if an edge of a particular type should exist
between those nodes.\footnote{When the graph corresponds to a knowledge base
like Freebase, this task is also known as knowledge base completion.}

We differ from the work done by Gardner and Mitchell in two ways.  First, we
apply these features to models of natural language, instead of to the link
prediction task.  Second, Gardner and Mitchell only considered generating
feature matrices over node \emph{pairs} in a graph, with which one can model
binary relations.  The natural language queries we wish to answer also involve
unary predicates.  We thus extend SFE to also generate feature matrices over
\emph{nodes} in a graph, a straightforward extension that is described in
\secref{method}.

\section{Method}
\label{sec:method}

This section presents our approach to open vocabulary semantic parsing. Our
approach follows the same structure as described in Section
\ref{sec:jayant-semparse}: we generate a surface logical form from a query,
then evaluate it against a probabilistic database with learned predicate
instance probabilities. Section \ref{sec:better-lfs} describes our logical form
generation, which improves the analyses of noun-mediated relations relative to
prior work. Sections \ref{sec:formal-and-distributional} and
\ref{sec:feature-generation} describe our predicate models and the
corresponding feature generation process. Finally, Section
\ref{sec:better-candidates} describes an improvement on candidate entity
generation for question answering.

\subsection{Logical form generation}
\label{sec:better-lfs}

As done by Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}, we generate
logical forms from natural language statements by computing a syntactic CCG
parse, then applying a collection of rules to produce logical forms. However,
their logical form analyses do not model noun-mediated relations well. For
example, given the phrase ``Italian architect Andrea Palladio,'' their system's
logical form would include the relation $\lexicalpredicate{N/N}(\entity{Italy},
\entity{Andrea Palladio})$. Here, the \lexicalpredicate{N/N} predicate
represents a generic noun modifier relation; however, this relation is too
vague for the predicate model to accurately learn its denotation. A similar
problem occurs with prepositions and possessives, e.g., it is similarly hard to
learn the denotation of the predicate \lexicalpredicate{of}.

Our system improves the analysis of noun-mediated relations by simply including
the noun in the predicate name. In the architect example above, our system
produces the relation \lexicalpredicate{architect\_N/N}. It does this by
concatenating all intervening noun modifiers between two entity mentions and
including them in the predicate name; for example, ``Illinois attorney general
Lisa Madigan'' produces the predicate
\lexicalpredicate{attorney\_general\_N/N}).\footnote{While it would be
preferable to use actual dependency relations to determine which words to add
to the predicate, this is not possible because our parser produces flat
analyses of noun phrases where all noun modifiers depend directly on the head
of the noun phrase.} We similarly improve the analyses of prepositions and
possessives to include the head noun. For example, ``Barack Obama, president of
the U.S.'' produces the predicate instance
$\lexicalpredicate{president\_of}(\entity{Barack Obama}, \entity{U.S.})$, and
``Rome, Italy's capital'' produces the predicate
\lexicalpredicate{'s\_capital}. This process generates more specific predicates
that we empirically found were easier for the predicate models to learn.

\subsection{Predicate models}
\label{sec:formal-and-distributional}

The key question in the semantic parsing framework we are using is how to
define the predicate models that underlie the probabilistic database described
in \secref{jayant-semparse}.  This is where previous work used purely
distributional models, and where we wish to include structured information from
a knowledge base to improve the semantics of these predicate models.

We inject information from a knowledge base into our predicate models by
augmenting the learned embeddings with features. In our models, the truth
probability of a category instance $c(e)$ or relation instance $r(e_1, e_2)$ is
given by:

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}

In the above equation, the $\theta$ and $\phi$ terms are learned predicate and
entity embeddings, as described in Section \ref{sec:jayant-semparse}. The
second term in the sum represents our new features and their learned weights.
$\psi_c(e)$ and $\psi_r(e_1, e_2)$ are SFE feature vectors for each entity and
entity pair; a different set of features is chosen for each predicate $c$ and
$r$ in order to maximize their usefulness. The terms $\omega_c$ and $\omega_r$
represent learned weights for these features.

Each SFE feature can be viewed as either a Freebase query or a horn clause
implication rule using the Freebase schema. In both cases, the feature defines
a set of entities (or entity pairs), and the associated feature weight measures
the likelihood that an entity in this set is also in the denotation of the
surface predicate. Our models include \emph{many} such features for each
surface predicate, effectively mapping each surface predicate onto a
\emph{weighted combination of Freebase queries}. This approach is fundamentally
different from typical semantic parsers that map language to a single Freebase
query.

In our model, there are now three sets of parameters to be learned: (1)
$\theta$, low-dimensional distributional vectors trained for each predicate;
(2) $\phi$, low-dimensional distributional vectors trained for each entity and
entity pair; and (3) $\omega$, weights associated with the selected formal SFE
features for each predicate.  All of these parameters are optimized jointly,
using the semantic parsing framework described in \secref{jayant-semparse}.

\subsection{Generating predicate features}
\label{sec:feature-generation}

The feature vectors $\psi_c(e)$ and $\psi_r(e_1, e_2)$ are constructed using a
two-stage process. First, we generate a predicate-independent feature vector
$\psi(e)$ for each entity $e$ and $\psi(e_1, e_2)$ for each entity pair $(e_1,
e_2)$ using SFE. This step generates very high-dimensional feature vectors,
where most of the features are not relevant for any given predicate. Next, we
perform per-predicate feature selection to identify a subset of these features
that are relevant for each category and relation predicate.

The predicate-independent feature extraction stage computes the following
feature vectors. For entity pairs, the vector $\psi(e_1, e_2)$ is a sparse
binary vector containing indicator features for all edge sequences that connect
nodes $e_1$ and $e_2$ in the Freebase graph, up to length 4.  For an entity
$e$, the vector $\psi(e)$ contains two kinds of indicator features: (1) all
edge sequences up to length two that begin at $e$, and (2) the same sequences
paired with the entity at their endpoint. Features of type (1) capture the
kinds of relations that an entity participates in, e.g., designing
architectural structures, while features of type (2) capture relations to other
particular entities. This second kind of feature is most helpful when the
endpoint entity is a type, such as \entity{architect}, although there are some
other valuable features of this kind. \figref{sfe} shows some example entity
and entity pair features extracted from a subset of the Freebase graph.

The feature selection step uses pointwise mutual information (PMI) to reduce
the number of features considered by the model for each predicate. The feature
vectors generated above have tens of millions of dimensions, but, for any given
surface predicate, only a handful of these dimensions represent relevant
Freebase queries. We select features by first summing the entity and entity
pair feature vectors seen with each predicate in the training data. For
example, the phrase ``Italian architect Andrea Palladio'' is considered a
positive training example for the instantiated predicates
$\formalpredicate{architect}(\entity{Andrea Palladio})$ and
$\formalpredicate{architect\_N/N}(\entity{Italy}, \entity{Andrea Palladio})$.
We then add the feature vectors for \entity{Andrea Palladio} and
(\entity{Italy}, \entity{Andrea Palladio}) to the feature counts for the
predicates \formalpredicate{architect} and \formalpredicate{architect\_N/N},
respectively. This gives a set of counts \formalpredicate{count}($\pi$),
\formalpredicate{count}($f$), and \formalpredicate{count}($\pi\land f$), for
each predicate $\pi$ and feature $f$.  The features are then ranked by PMI for
each feature by computing $\frac{\formalpredicate{count}(\pi\land
f)}{\formalpredicate{count}(\pi)\formalpredicate{count}(f)}$.  After removing
features with counts below a threshold, we pick the $k=100$ features with the
highest PMI values for each predicate to use in our model.

\subsection{Candidate entity generation}
\label{sec:better-candidates}

A key benefit of our predicate models is that they are able to assign scores to
entity pairs that were never seen in the training data. The purely
distributional model of Krishnamurthy and Mitchell has no basis for assigning
these scores and therefore assumes $P(r(e_1,e_2)) = 0$ for unseen entity pairs
$(e_1,e_2)$. This assumption limits the recall of their system when it is
applied to question answering, as these entity pairs will not have been
observed for many correct, but rare entity answers. In contrast, our predicate
models can score any entity pair: the distributional component of an unseen
pair will be zero, but the SFE component is always applicable. This benefit
allows our model to considerably improve question answering performance on rare
entities.

However, this benefit has an associated drawback: given a logical form query,
it is computationally and statistically undesirable to consider \emph{all}
Freebase entities as answers. Scoring millions of entities is slow, and
furthermore makes it more difficult to identify correct answers (as their
relative frequency in the candidate set decreases).  Therefore, when answering
a query, we compute a set of candidate entities and only score these
candidates. Our candidate set consists of all entities seen with the query
entities in the training set as well as all entities directly connected to the
query entities in Freebase, or connected by a mediator node.\footnote{Mediators
in Freebase are used to capture relations with more than two arguments, such as
employment tenure, which has an employer, and employee, a start date, and an
end date.}

Unfortunately, for some entities, such as the United States, the set of
connected entities can be intractably large; we therefore limit this expansion
to only those entities with fewer than 100 directly connected entities.  Our
motivation for this is that this candidate entity generation is most useful for
rarely seen entities, for which we have few or no related entities seen during
training.  These entities also tend to have relatively few connected entities
in Freebase.

\section{Evaluation}
\label{sec:evaluation}

We evaluate our open-vocabulary semantic parser on a fill-in-the-blank natural
language query task.  Each test example is a natural language phrase containing
at least two Freebase entities, one of which is held out.  The system must
propose a ranked list of Freebase entities to fill in the blank left by the
held out entity, and the predicted entities are then judged manually for
correctness.  We compare our proposed models, which combine distributional and
formal elements, with a purely distributional baseline from prior work.  All of
the data and code used in these experiments is available at [url withheld for
review].

\subsection{Data}

We use the dataset introduced by Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}, which
consists of the ClueWeb09 web
corpus\footnote{http://www.lemuproject.org/clueweb09.php} along with Google's
FACC entity linking of that corpus to
Freebase~\cite{gabrilovich-2013-clueweb-entity-linking}.  For training data, 3
million webpages from this corpus were processed with a CCG parser to produce
logical forms~\cite{krishnamurthy-2014-joint-ccg}.  This produced 2.1m
predicate instances involving 142k entity pairs and 184k entities.  After
removing infrequently-seen predicates (seen fewer than 6 times), there were 25k
categories and 4.2k relations.\footnote{The differences in numbers reported
here versus those reported by Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} are due to our
improved logical form generation, discussed in \secref{better-lfs}.}

We also used the test set created by Krishnamurthy and Mitchell, which contains
220 queries generated in the same fashion as the training data from a separate
section of ClueWeb.  However, as they did not release a development set with
their data, we used this set as a development set.  For a final evaluation, we
generated another, similar test set from a different held out section of
ClueWeb, in the same fashion as done by Krishnamurthy and Mitchell.  This final
test set contains 307 queries.  We report results on both of these sets below.

\subsection{Models}

We compare three models in our experiments: (1) a distributional model (the
prior work of Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}); (2) a formal
model (new to this work), where the distributional parameters $\theta$ and
$\phi$ in \secref{formal-and-distributional} are fixed at zero; and (3) the
combined model described in \secref{formal-and-distributional} (also new to
this work).  Except where noted, all experiments use our modified logical forms
(\secref{better-lfs}) and our entity proposal mechanism
(\secref{better-candidates}).

\subsection{Methodology}

Given a fill-in-the-blank query such as ``Italian architect \blank{}'', each
system produces a ranked list of 100 candidate entities.  To compare the output
of the systems, we follow a pooled evaluation protocol commonly used in
relation extraction and information
retrieval~\cite{west-2014-kbc-via-qa,riedel-2013-mf-universal-schema}.  We take
the top 30 predictions from each system and manually annotate whether they are
correct, and use those annotations to compute the average precision (AP) and
reciprocal rank (RR) of each system on the query.  Average precision is defined
as $\frac{1}{m}\sum^m_{k=1} \mathrm{Prec}(k) \times \mathrm{Correct}(k)$, where
$\mathrm{Prec}(k)$ is the precision at rank $k$, $\mathrm{Correct}(k)$ is an
indicator function for whether the $k$th answer is correct, and $m$ is number
of returned answers (up to 100 in this evaluation).  AP is equivalent to
calculating the area under a precision-recall curve.  Reciprocal rank is
computed by first finding the rank $r$ of the first correct prediction made by
a system.  Reciprocal rank is then $\frac{1}{r}$, ranging from 1 (if the first
prediction is correct) to 0 (if there is no correct answer returned).  In the
tables below we report \emph{mean} average precision (MAP) and \emph{mean}
reciprocal rank (MRR), averaged over all of the queries in the test set.  We
also report a weighted version of MAP, where the AP of each query is scaled by
the number of annotated correct answers to the query (shown as W-MAP in the
tables for space considerations).

\subsection{Results}

\begin{table}
  \centering
  {\small
    \begin{tabular}{lcc}
      \toprule
      Method & K\&M's LFs & Our LFs \\
      \midrule
      Distributional model & .269 & \textbf{.284} \\
      \midrule
      Formal model & .231 & \textbf{.276} \\
      \midrule
      Combined model & .313 & \textbf{.335} \\
      \bottomrule
    \end{tabular}
  }
  \caption{Comparison of models using logical forms from prior work versus
  those introduced in this paper.  The metric shown is mean average precision
  on the dev set.  With our logical forms, performance of all models improves,
  quite substantially in the case of the formal model.}
  \label{tab:better-lfs}
\end{table}

We first show the effect of using the new logical forms introduced in
\secref{better-lfs}.  As can be seen in \tabref{better-lfs}, with improved
logical forms, all models are better able to capture the semantics of language.
This improvement is more pronounced in the formal models, which have more
capacity to get specific features from Freebase with the new logical forms.  As
our logical forms are able to give all models better performance, the remaining
experiments we present all use these logical forms.

\begin{table}
  \centering
  {\small
    \begin{tabular}{lccc}
      \toprule
      Method & MAP & W-MAP & MRR \\
      \midrule
      Distributional model & .163 & .163 & .288 \\
      \midrule
      With freebase entities & .\textbf{229} & \textbf{.275} & \textbf{.312} \\
      \midrule
      \midrule
      Relative improvement & 40\% & 69\% & 8\% \\
      \bottomrule
    \end{tabular}
  }
  \caption{Allowing Krishnamurthy and Mitchell's model to propose candidate
  entities from Freebase improves mean average precision substantially, by
  allowing the model to have higher recall.}
  \label{tab:better-candidates}
\end{table}

We next show the improvement gained by using the simple candidate entity
generation outlined in \secref{better-candidates}.  By simply appending the
list of connected entities in Freebase to the end of the rankings returned by
the distributional model, MAP improves by 40\% (see
\tabref{better-candidates}).  The connectedness of an entity pair in Freebase
is very informative, especially for rare entities that are not seen together
during training.

\begin{table}
  \centering
  {\small
    \begin{tabular}{lccc}
      \toprule
      Method & MAP & W-MAP & MRR \\
      \midrule
      Prior work (distributional) & .284 & .371 & .379 \\
      \midrule
      Our formal model & .276 & .469 & .334 \\
      \midrule
      Our combined model & \textbf{.335} & \textbf{.477} & \textbf{.429} \\
      \midrule
      \midrule
      Relative improvement & 18\% & 29\% & 13\% \\
      \bottomrule
    \end{tabular}
  }
  \caption{Results on the development set for our fill-in-the-blank task.  The
  combined model significantly improves MAP over prior work.}
  \label{tab:dev-results}
\end{table}

\tabref{dev-results} shows a comparison between the semantic parsing models we
have discussed on the development set.  As can be seen, the combined model
significantly improves performance over prior work, giving a relative gain in
weighted MAP of 29\%.

\begin{table}
  \centering
  {\small
    \begin{tabular}{lccc}
      \toprule
      Method & MAP & W-MAP & MRR \\
      \midrule
      Prior work (distributional) & .229 & .275 & .312 \\
      \midrule
      Our formal model & .355 & .495 & .419 \\
      \midrule
      Our combined model & \textbf{.370} & \textbf{.513} & \textbf{.469} \\
      \midrule
      \midrule
      Relative improvement & 62\% & 87\% & 50\% \\
      \bottomrule
    \end{tabular}
  }
  \caption{Results on the final test set for our fill-in-the-blank task.  The
  combined model improves over prior work by 50--87\% on our metrics.  Note
  that these improvements over the baseline are \emph{after} the baseline has
  been improved by the methods developed in this paper, shown in
  \tabref{better-lfs} and \tabref{better-candidates}.  The cumulative effect of
  the methods presented in this work is an improvement of over 120\% in MAP.}
  \label{tab:final-results}
\end{table}

\tabref{final-results} shows that these improvements are consistent on the
final test set, as well.  The performance improvement seen by the combined
model is actually larger on this set, with gains on our metrics ranging from
50\% to 87\%.

On both of these datasets, the difference in MAP between the combined model and
the distributional model is statistically significant (by a paired permutation
test, $p < 0.05$).  The differences between the combined model and the formal
model, and between the formal model and the distributional model, are not
statistically significant, as each method has certain kinds of queries that it
performs well on.  Only the combined model is able to consistently outperform
the distributional model on all kinds of queries.

\subsection{Discussion}

Our model tends to outperform the distributional model on queries containing
predicates with exact or partial correlates in Freebase. For example, our model
obtains nearly perfect average precision on the queries ``French newspaper
\blank{}'' and ``Israeli prime minister \blank{},'' both of which can be
exactly expressed in Freebase.  The top features for
\lexicalpredicate{newspaper}($x$) all indicate that $x$ has type
\formalpredicate{newspaper} in Freebase, and the top features for
\lexicalpredicate{newspaper\_N/N}($x$, $y$) indicate that $y$ is a newspaper,
and that $x$ is either the circulation area of $y$ or the language of $y$.

The model also performs well on queries with partial Freebase correlates, such
as ``Microsoft head honcho \blank{}'', ``The United States, \blank{}'s closest
ally'', and ``Patriots linebacker \blank{},'' although with somewhat lower
average precision. The high weight features in these cases tend to provide
useful hints, even though there is no direct correlate; for example, the model
learns that ``honchos'' are people, and that they tend to be CEOs and film
producers.

There are also some areas where our model can be improved. First, in some
cases, the edge sequence features used by the model are not expressive enough
to identify the correct relation in Freebase. An example of this problem is the
``linebacker'' example above, where the features for
\lexicalpredicate{linebacker\_N/N} can capture which athletes play for which
teams, but not the \emph{positions} of those athletes. Second, our model can
underperform on predicates with no close mapping to Freebase. An example where
this problem occurs is the query ``\blank{} is a NASA mission.'' Third, there
remains room to further improve the logical forms produced by the semantic
parser, specifically for multi-word expressions. One problem occurs with
multi-word noun modifiers, e.g., ``Vice president Al Gore'' is mapped to
$\lexicalpredicate{vice}(\entity{Al Gore}) \land
\lexicalpredicate{president}(\entity{Al Gore})$. Another problem is that there
is no backoff with multi-word relations. For example, the predicate
\lexicalpredicate{head\_honcho\_N/N} was never seen in the training data, so it
is replaced with \lexicalpredicate{unknown}; however, it would be better to
replace it with \lexicalpredicate{honcho\_N/N}, which \emph{was} seen in the
training data. Finally, although using connected entities in Freebase as
additional candidates during inference is helpful, it often over- or
under-generates candidates. A more tailored, per-query search process could
improve performance.

\section{Related work}

In addition to the work described in \secref{background}, there are three broad
classes of work that are related to what we have presented in this paper.

The first line of research that our work is related to is knowledge base
completion.  In addition to SFE~\cite{gardner-2015-sfe}, our work draws
inspiration from work on embedding the entities and relations in a knowledge
base~\cite{riedel-2013-mf-universal-schema,%
nickel-2011-rescal,bordes-2013-transe,nickel-2014-are,%
toutanova-2015-joint-text-kb-embedding}, as well as work on graph-based methods
for reasoning with knowledge
bases~\cite{lao-2010-original-pra,gardner-2014-vector-space-pra,%
neelakantan-2015-rnn-kbc}.

Second, our work is conceptually related to many methods that aim to learn word
embeddings that are informed by some kind of external
knowledge~\cite{faruqui-2015-retrofitting-word-vectors,%
rocktaschel-2015-logical-embeddings,schwartz-2016-symmetric-patterns-w2v,%
yu-2014-w2v-with-semantic-knowledge}.  A key difference between these
approaches and ours is that they use the structured information only to modify
distributional representations, while we include the structured information
directly in our models, training feature weights jointly with distributional
representations.

Lastly, there is an extensive literature on building semantic parsers to answer
questions against a structured knowledge
base~\cite{zettlemoyer-2005-ccg,berant-2013-semantic-parsing-qa,%
kwiatkowski-2013-ontology-matching,krishnamurthy-2012-semantic-parsing,%
li-2015-semantic-parsing-scfg}.  The vast majority of this work is focused on
matching natural language text to statements in a specific schema, with no hope
of assigning meaning to language that falls outside of that schema.  Only very
recently has there been work attempting to learn broad coverage meaning
representations in semantic parsers.  In addition to the work by Krishnamurthy
and Mitchell discussed in \secref{background}, and work by Lewis and
Steedman~\shortcite{lewis-2013-combined-distributional-and-logical-semantics}
which Krishnamurthy and Mitchell extended, Choi et
al.~\shortcite{choi-2015-semantic-parsing-partial-ontologies}, extended a
traditional semantic parser with the notion of ``open'' predicates not
contained in the Freebase ontology.  A key difference between their work and
ours is that, while their system might learn that ``front-runner'' should be
modeled as an open predicate that has no representation in Freebase, they have
no method to assign any further meaning to that predicate, whereas our models
learn both distributional and formal information about the meaning of
``front-runner''.

\section{Conclusion}
\label{sec:conclusion}

This paper considers the problem of assigning meaning to predicates (both
categories and relations) in natural language using both distributional
information and information from a formal knowledge base. Our model represents
words as a weighted combination of knowledge base queries plus a distributional
component. Both of these components are jointly trained to correctly predict
entity answers to fill-in-the-blank natural language queries. This combined
model achieved relative gains of over 50\% in mean average precision and mean
reciprocal rank versus a purely distributional approach.  We also introduced a
better mapping from surface text to logical forms, and a simple method for
using Freebase to find candidate entities during inference.  Taken together,
the methods introduced in this paper improved mean average precision on our
task from .163 to .370, a 127\% relative improvement over prior work.

A consequence of this work is that it suggests a new direction for semantic
parsing research. Existing semantic parsers map language to a single knowledge
base query, an approach that successfully leverages a knowledge base's
predicate instances, but is fundamentally limited by its schema. In contrast,
our approach maps language to a \emph{weighted combination of queries} plus a
distributional component; this approach is capable of representing a much
broader class of concepts while still using the knowledge base when it is
helpful. Furthermore, it is capable of using the knowledge base even when the
meaning of the language cannot be exactly represented by a knowledge base
predicate, which is a common occurrence. We believe that this kind of approach
could significantly expand the applicability of semantic parsing techniques to
more complex domains where the assumptions of traditional techniques are too
limiting.

%\section*{Acknowledgments}

\bibliography{bib}
\bibliographystyle{acl2016}

\end{document}